### 2014-03-11 paper.Rnw

% !Rnw weave = knitr
\documentclass{article}

\usepackage{cite}
\begin{document}

<<setup, include=FALSE, cache=FALSE, echo=FALSE>>=
Sys.setenv(TEXINPUTS=getwd(), 
           BIBINPUTS=getwd(), 
           BSTINPUTS=getwd())
@

\title{Analysis of Massachusetts Standardized Testing Data through Structural Equation Modelling}
\date{March 24, 2014}
\author{Landon Hurley\\ Psychology, SUNY Geneseo \and Herb Susmann\\ Mathematics, SUNY Geneseo}

\maketitle

\section*{Abstract}

\section*{Introduction}

\section*{Data}

The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998. We examine a random sample of 10,515 10th grade student's results from the 2009 MCAS test. The data also contains basic demographic information for each student.

The test consists of english, mathematics and science components. Every student takes a English and Mathematics tests; the science component can be fulfilled by taking either a Biology, Chemistry, Introductory Physics, or Technology/Engineering test. \cite{mcas_summary}

Each test consists of multiple choice and short answer type questions.

\section*{Methods}
Description of SEM

\section*{Results}

\section*{Discussion}
Could also include ideas for other questions to ask in the questionairre

\section*{Works Cited}

\bibliography{references}{}
\bibliographystyle{plain}

\end{document}

### 2014-03-11 EDA work.R
MI.model syntax and files fixed.

@@ -203,7 +203,7 @@ d2<-d[,3:90]
 # d2[,44:46]<-list(NULL)
 codive <- function(x) {ifelse(x=='+',1,ifelse(x==' ',NA,0))}
 d3<-data.frame(codive(d2))
-d3<-as.data.frame(lapply(d3,factor))
+# d3<-as.data.frame(lapply(d3,factor))
 d3$eitem9<-d2$eitem9
 d3$eitem18<-d$eitem18
 d3$eitem27<-d$eitem27
@@ -223,10 +223,11 @@ d3$mrawsc<-d$mrawsc
 d3$eitem41<-NULL
 d3$eitem42<-NULL
 
-#d3<-d3[d3$eth!='M' & d3$eth!='P' & d3$eth!='N' & d3$quest3!='E',]
-d3<-d3[d3$eth!='M' & d3$eth!='P' & d3$eth!='N' & d3$quest3!='E',]
+#d3<-d3[d3$eth!='M' & d3$eth!='P' & d3$eth!='N',]
+d3$eth<-d$eth
+d3<-d3[d3$eth!='M' & d3$eth!='P' & d3$eth!='N',]
+require(gdata)
 d3$eth<-drop.levels(d3$eth,reorder=T)
-
 
 invariance="f1=~eitem8+eitem13+eitem22+eitem26+eitem28+eitem30+eitem31
             f2=~eitem38+eitem39+eitem40
@@ -266,4 +267,4 @@ sci1$eth<-da$eth
 sci1<-sci1[sci1$eth !='M' & sci1$eth !='P' & sci1$eth !='N',]
 sci1$eth<-drop.levels(sci1$eth,reorder=T)
 
-MI.model2<-measurementInvariance(sci_inv,data=sci1,group='eth',missing='fiml',bootstrap=75)
+MI.model2<-measurementInvariance(sci_inv,data=sci1,group='eth',missing='fiml',bootstrap=100)

### 2014-03-13 paper.Rnw
Notes on direction for paper

@@
 
+- Questionairre results - regression tree
+
+Results from "omega" exploratory data analysis
+- diagram of factor loadings
+
+Model Structure
+
+Table
+- Results from factor invariance
+
 \section*{Discussion}
+Test doesn't work the same across races
+
+Could be because of SES?
+
+Questionairre results
+
 Could also include ideas for other questions to ask in the questionairre

### 2014-03-17 EDA work.R
something changed, can't remember what. edit: Maybe it has the manova/mvpart results which find no relation between questions/demos and scores.

@@ -238,7 +238,9 @@ invariance="f1=~eitem8+eitem13+eitem22+eitem26+eitem28+eitem30+eitem31
             erawsc~f1+f2
             mrawsc~f4+f5+f6+f7"
 
-MI.model<-measurementInvariance(invariance,data=d3,group='eth',missing='fiml',bootstrap=75)
+MI.model<-measurementInvariance(invariance,data=d3,group='eth',missing='fiml',bootstrap=100)
+
+
 
 sci_inv<-'f3=~sitem1+sitem2+sitem13+sitem14+sitem15+sitem22+sitem27+sitem30+sitem33+sitem34+sitem37+sitem40+sitem41
 f2=~sitem3+sitem5+sitem18+sitem19+sitem28+sitem31+sitem35+sitem43+sitem45
@@ -267,4 +269,17 @@ sci1$eth<-da$eth
 sci1<-sci1[sci1$eth !='M' & sci1$eth !='P' & sci1$eth !='N',]
 sci1$eth<-drop.levels(sci1$eth,reorder=T)
 
-MI.model2<-measurementInvariance(sci_inv,data=sci1,group='eth',missing='fiml',bootstrap=100)
+MI.model2<-measurementInvariance(sci_inv,data=sci1,group='eth',missing='fiml',bootstrap=100,strict=T,test='Bollen.Stine')
+
+
+d4<-d[,145:172]
+d5<-d[,c(170,141,91,45)]
+d4<-merge(d4,d5,by='id')
+require(mvpart)
+d4$id<-NULL
+fit1<-mvpart(cbind(d$mrawsc,d$erawsc,d$srawsc)~sex+eth,data=d4,)
+
+summary(manova(cbind(d$mrawsc,d$erawsc,d$srawsc)~.,d4))
+
+
+

### 2014-03-18 paper.Rnw
rewrote everything. also screwed up the manova and tree results, pulling variables across two data.frames.

--- 2014-03-11 paper.Rnw
@@
-% !Rnw weave = knitr
 \documentclass{article}
-
-\usepackage{cite}
+\usepackage{apacite,geometry,fancyhdr}
+\pagestyle{fancy}
+\rhead{Hurley \& Susmann}
+\lhead{\thepage}
 \begin{document}
 
-<<setup, include=FALSE, cache=FALSE, echo=FALSE>>=
+<<setup, include=T, cache=FALSE, echo=FALSE>>=
 Sys.setenv(TEXINPUTS=getwd(), 
@@
 
-\title{Analysis of Massachusetts Standardized Testing Data through Structural Equation Modelling}
+\title{An Analysis of Massachusetts' Standardized Testing through Multi-Group Structural Equation Modelling}
 \date{March 24, 2014}
@@
 
-\section*{Abstract}
+\section{Abstract}
 
-\section*{Introduction}
+\section{Introduction}
 
-\section*{Data}
+\subsection{Data}
 
 The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
-We examine a random sample of 10,515 10th grade student's results from the 2009 MCAS test.
-The data also contains basic demographic information for each student.
+We examine a subset sample of 10\textsuperscript{th} grade students' results from the 2009 Spring MCAS test.
+The data also contains basic demographic information for each student comprised of race/ethnicity, gender, and an academic engagement questionnaire.
+Each test consists of multiple choice items and open ended type questions that are graded on a holistic scale, implemented as a 5 point integer scale.
 
-The test consists of english, mathematics and science components.
-Every student takes a English and Mathematics tests; the science component can be fulfilled by taking either a Biology, Chemistry, Introductory Physics, or Technology/Engineering test.
+The examinations are comprised of three separate knowledge domains: English, Mathematics and a science component.
+Every student takes identical versions of the English and Mathematics sections; the science component can be fulfilled by taking either a Biology, Chemistry, Introductory Physics, or Technology/Engineering test.
 \cite{mcas_summary}
 
-Each test consists of multiple choice and short answer type questions.
+The MCAS examination structure has received substantial criticism from both educational activists and theoreticians.
+The former, comprised mainly of teachers and educational policy analysts, claim that the test is unfairly biased against students as a function of demographics, while citing the significant impediment that poor performance upon the tests has upon both immediate and long-term outcomes.
+Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
 
-\section*{Methods}
-Description of SEM
+Our analysis begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into  performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+\section{Methods}
+\subsection{Subjects}
+Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusett's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} grade ethnicity characteristics as 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female.
 
-\section*{Results}
+Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relaionship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
+\subsection{Procedure}
+A critical assumption within standardised testing is that the test demonstrates a psychometric property called measurement invariance/equivalence.
+This property is a series of mathematical constraints applied to latent variables.
+Latent variables are unobservable constructs measured in an identical process to standard principal axis factor analysis.
+The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample, by imposing restrictions in a iterative four phase procedure: \begin{itemize}
+\item{Configural invariance:} 
+Tests whether the same factor model (i.e., latent variables, equivalent to knowledge subdomains, for example grammar) is found within each subgroup.
+\item{Weak invariance:}
+Tests whether the same items load equivalently onto the same same structure, presuming that configural invariance has been demonstrated.
+\item{Strong invariance}
+Tests that the intercepts are equivalent across groups.
+\item{Strict invariance:}
+Tests the assumption that residual variances are equivalent across groups.
+\end{itemize}
+These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
+In this paper, we examine these factors using a series of R packages found cited at the end of the paper \cite{psych,lavaan,semTools}.
+Exploratory bootstrapped oblimin factor analyses were conducted upon each section to establish the number of domains and high loading items within each of the three tests.
+As a consequence of the extreme ethnicity imbalance within three of the four science exams, only Biology was tested using MG-CFA.
 
-\section*{Discussion}
-Could also include ideas for other questions to ask in the questionairre
+Furthermore, we investigated the information found within the questionnaire, relating them to the students' scores on the three sections using both multivariate parametric regressions, and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
 
-\section*{Works Cited}
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
+Even so, for the MG-CFA, we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation, in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques, producing potentially biased results under the assumption that items are missing completely at random.
+As such, we processed the original responses using the mice package\cite{mice}, using 25 imputed data frames with 30 iterations in each dataframe.
+Regression trees are relatively robust to missing data, because all information is not processed simultaneously, but sequentially in order of variable importance, and as such we did not take any additional steps when conducting that analysis.
+\section{Results}
+\subsection{Theoretical structure of the test}
+Substantive theoretical knowledge of the number of latent knowledge domains that were intended to be tested was generated from the answer keys provided by Massachusetts' Department of Education.
+Specifically, there were 5 general topics within Biology:
+\begin{itemize}
+\item{Anatomy \& Physiology}
+\item{Biochemistry and Cell Biology}
+\item{Ecology}
+\item{Evolution and Biodiversity}
+\item{Genetics}
+\end{itemize}
+Two topics within English:
+\begin{itemize}
+\item{Reading \& Literature}
+\item{Language}
+\end{itemize}
+and five topics within Mathematics:
+\begin{itemize}
+\item{Data Analysis  Statistics  and Probability}
+\item{Geometry}
+\item{Measurement}
+\item{Number Sense and Operations}
+\item{Patterns Relations and Algebra}
+\end{itemize}
+\subsection{Exploratory factor analyses}
+With the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains, and factor loadings calculated using maximum likelihood estimates calculated using the tetrachoric covariance matrices for each subset of items.
+The psych package was utilised both to calculate the factor structure.
+Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
+\subsection{MG-CFA}
+As a consequence of the unequal numbers of test takers between Biology and the rest of the exam, we chose to run two separate tests for measurement invariance by running Biology separately.
+Using the results of the EFA, we specified the following two models:
+<<>>=
+invariance
+sci_inv
+@
+These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+Others were excluded due to sample size and composition constraints.
+As a consequence of the $\chi$\textsuperscript{2} fit statistic scaling as a function of both sample size and the minimum function test statistic, traditional measurement invariance techniques of using changes in significance to reject a constraint fail here, as the sample size guarantees significant results.
+In response to this fact, we employ two different techniques: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
+For both, decreases in model fit of greater than .01 are considered significant.
+In addition, we include An (Akaike's) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps.
+
+\subsection{English and Mathematics}
+\subsubsection{Configural Invariance}
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
+
+<<>>=
+fitMeasures(MI.model$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+\subsubsection{Metric Invariance}
+
+<<>>=
+fitMeasures(MI.model$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+The second test demonstrates a failure of metric/weak invariance when comparing either the $\Delta$ TLI or $\Delta$ CFI.
+Commonly, metric invariance is used to demonstrate that for all items, one unit change in the item scores is scaled to an equal unit change in the factor score across groups.
+This implies that the variances on each measure are the same for all items.
+As a consequence of the failure of metric invarianc to hold, all subsequent equivalences are rejected as well.
+
+\subsection{Biology}
+We proceeded with the same methodology to test the science section in isolation, using identical $\Delta$ constraints as above.
+The test for configural invariance produced non-significant results, indicating general acceptance of the factor structure.
+In addition, each subsequent test produced similar results, indicating that across all four primary ethnicities, the test operated with equivalent measurement.
+
+<<>>=
+fitMeasures(MI.model2$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.model2$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.model2$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.model2$fit.means, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+\subsection{Questionnaire}
+
+<<>>=
+require(mvpart)
+fit1<-mvpart(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,data=d4)
+summary(manova(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
+summary(lm(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
+@
+
+\section{Discussion}
+As indicated in the results section, the test fails to perform equivalently across the four groups, which is the primary point of order in test design and advocacy.
+This lends substantive credence to the argument that there are fundamental flaws within the test.
+However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
+In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between the general education practices taught, which is likely a function of SES, which disproportionately affects minorities over Whites.
+This subtle distinction has the immediate consequence of arguing that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
+
+Little more can be said about the questionnaire beyond the fact that the mode for each item is the most efficient predictor of individual student responses.
+As such, we can only conclude that the information being measured by these items are independent of the exams' item functioning and student performance.
 
 \bibliography{references}{}
-\bibliographystyle{plain}
+\bibliographystyle{apacite}
 

### 2014-03-18 paper.tex
rewrote everything. also screwed up the manova and tree results, pulling variables across two data.frames.

--- 2014-03-18 paper.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{apacite,geometry,fancyhdr}
@@
 \lhead{\thepage}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<setup, include=T, cache=FALSE, echo=FALSE>>=
-Sys.setenv(TEXINPUTS=getwd(), 
-           BIBINPUTS=getwd(), 
-           BSTINPUTS=getwd())
-@
 
@@
 Using the results of the EFA, we specified the following two models:
-<<>>=
-invariance
-sci_inv
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{invariance}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'invariance' not found}}\begin{alltt}
+\hlstd{sci_inv}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'sci\_inv' not found}}\end{kframe}
+\end{knitrout}
+
 These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
@@
 
-<<>>=
-fitMeasures(MI.model$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<>>=
-fitMeasures(MI.model$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<>>=
-fitMeasures(MI.model2$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.means, c('chisq','tli', 'cfi','rmsea','aic'))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.means,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
 
 \subsection{Questionnaire}
+Preliminary results were conducted using a multivariate regression tree, which searches for both linear and non-linear relationships between different classifications of the three raw scores and the questionnaire items, ethncity, and gender.
+However, the data failed to split at all, meaning that there is little information that relates between the questionnaire, demographics, and individuals' scores.
+Likewise mutlivariate regression results included demonstrate the no meaningful results were identified, with the few significant results conceptually meaningless.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{require}\hlstd{(mvpart)}
+\end{alltt}
 
-<<>>=
-require(mvpart)
-fit1<-mvpart(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,data=d4)
-summary(manova(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
-summary(lm(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
-@
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: mvpart}}\begin{alltt}
+\hlstd{fit1} \hlkwb{<-} \hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(d}\hlopt{$}\hlstd{mrawsc, d}\hlopt{$}\hlstd{erawsc, d}\hlopt{$}\hlstd{srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'd4' not found}}\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{manova}\hlstd{(}\hlkwd{cbind}\hlstd{(d}\hlopt{$}\hlstd{mrawsc, d}\hlopt{$}\hlstd{erawsc, d}\hlopt{$}\hlstd{srawsc)} \hlopt{~} \hlstd{., d4))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'd4' not found}}\end{kframe}
+\end{knitrout}
 

### 2014-03-18 EDA work.R

@@ -3,226 +3,106 @@ d<-read.csv('~/Documents/workspace/DataCompete/Student Data Competition/MCAS.csv
 # a<-read.spss('~/Documents/workspace/DataCompete/Student Data Competition/missingDataCoded.sav',use.value.labels=T,to.data.frame=T,use.missings=T)
 colnames(d)<-tolower(colnames(d))
 d$id<-rownames(d)
-d$id<-factor(d$id)
-
-codive <- function(x) {ifelse(x=='+',1,ifelse(x==' ',NA,0))}
-
-poi<-as.data.frame(codive(d[sapply(d,is.factor)]))
-poi$gender<-poi$race_off<-NULL
-poi[,111:135]<-list(NULL)
-
+d$id<-as.integer(d$id)
 
 d$firstlanguage<-factor(d$firstlanguage)
 d$grade<-factor(d$grade)
 d$sex<-d$gender
 d$eth<-d$race_off
+d$race_off<-d$gender<-NULL
+a1<-as.data.frame(codive(a[sapply(a,is.factor)]))
+codive <- function(x) {ifelse(x=='+',1,ifelse(x==' ',NA,0))}
 
-d$race_off<-NULL
-d$gender<-NULL
+poi<-as.data.frame(cbind(codive(d[sapply(d,is.factor)]),d$id,d$scitry))
+poi$sex<-poi$eth<-poi$grade<-poi$firstlanguage<-NULL
+poi$id<-poi$V140
+poi$scitry<-poi$V141
+poi$V140<-poi$V141<-NULL
+poi[111:135]<-list(NULL)
 
-# unique(sort(d$firstlanguage))
-# cor(d[sapply(d,is.numeric)],use='pairwise.complete.obs')
-require(psych)
-require(GPArotation)
-# homals(poi[,1:110],level='n',ndim=3,rank=3)
-poe<-poi[,1:36]
-pom<-poi[,37:68]
-pos<-poi[,69:110]
-
-#composes correlated factors from domains' multiple choice questions
-fe<-fa.poly(poe,nfactors=2,n.iter=100,rotate='oblimin',fm='ml',oblique.scores=T)
[…]

### 2014-03-20 paper.Rnw
fixed final dataframes. All EFA and CFA code should work. text is incomplete atm. still need to find out how to plot means of each tree node, if possible. also, R code does not run in paper.rnw txt: Need to fix

@@
 
-Our analysis begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into  performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+\subsection{Theoretical structure of the test}
+Our paper begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 \section{Methods}
 \subsection{Subjects}
-Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusett's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} grade ethnicity characteristics as 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female.
+Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusett's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} graders.
+Provided demographics present as follows 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female, and there are no significant interactions between gender and ethnicity.
 
@@
 \subsection{Procedure}
-A critical assumption within standardised testing is that the test demonstrates a psychometric property called measurement invariance/equivalence.
-This property is a series of mathematical constraints applied to latent variables.
-Latent variables are unobservable constructs measured in an identical process to standard principal axis factor analysis.
-The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample, by imposing restrictions in a iterative four phase procedure: \begin{itemize}
+A critical assumption within standardised testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
+This property is a series of mathematical constraints applied to latent variables that are assumed to be non-significantly different between groups.
+Latent variables are unobservable constructs that underly the observed measures; one typically estiamtes them by computing standard principal axis factor analysis.
+The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample, by imposing restrictions in a sequential four phase procedure:
+\begin{enumerate}
 \item{Configural invariance:} 
@@
 \item{Weak invariance:}
-Tests whether the same items load equivalently onto the same same structure, presuming that configural invariance has been demonstrated.
+Tests whether in addition to the same factor structure, the same items load equivalently onto the same same structure.
 \item{Strong invariance}
-Tests that the intercepts are equivalent across groups.
+Tests that the intercepts, item loadings, and factor structure are equivalent across groups.
 \item{Strict invariance:}
-Tests the assumption that residual variances are equivalent across groups.
-\end{itemize}
+Tests the assumption that in addition to the preceeding steps, the residual variances are equivalent across groups.
+\end{enumerate}
 These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
@@
 \section{Results}
-\subsection{Theoretical structure of the test}
-Substantive theoretical knowledge of the number of latent knowledge domains that were intended to be tested was generated from the answer keys provided by Massachusetts' Department of Education.
-Specifically, there were 5 general topics within Biology:
-\begin{itemize}
-\item{Anatomy \& Physiology}
-\item{Biochemistry and Cell Biology}
-\item{Ecology}
-\item{Evolution and Biodiversity}
-\item{Genetics}
-\end{itemize}
-Two topics within English:
-\begin{itemize}
-\item{Reading \& Literature}
-\item{Language}
-\end{itemize}
-and five topics within Mathematics:
-\begin{itemize}
-\item{Data Analysis  Statistics  and Probability}
-\item{Geometry}
-\item{Measurement}
-\item{Number Sense and Operations}
-\item{Patterns Relations and Algebra}
-\end{itemize}
 \subsection{Exploratory factor analyses}
-With the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains, and factor loadings calculated using maximum likelihood estimates calculated using the tetrachoric covariance matrices for each subset of items.
-The psych package was utilised both to calculate the factor structure.
+Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
+The psych package was used to calculate the factor structure.
 Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
 \subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the exam, we chose to run two separate tests for measurement invariance by running Biology separately.
-Using the results of the EFA, we specified the following two models:
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
+<<fig=T>>=
+fa.diagram(feng)
+@
+
+<<fig=T>>=
+fa.diagram(fmat)
+@
+
+<<fig=T>>=
+fa.diagram(fsci1)
+@
+
+These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+Other ethnicities were excluded because of their relative rarity within the sample.
+As a consequence of the $\chi$\textsuperscript{2} fit statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0 \end{equation} scales as a function of both sample size and the minimum function test statistic $F_0$, traditional measurement invariance, which uses changes in significance to reject a constraint, fails here, as the sample size guarantees significant results.
+In response to this fact, we employ three different fitness criterion: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
+For all three, decreases in model fit of greater than .01 are considered significant reductions in model fit.
+In addition, we include An (Akaike) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps in comparison to the theoretical true model.
+
+\subsection{English}
+The English section constructed using the following system of latent and test items, as predictors of the unstandardized English score, as given below.
 <<>>=
-invariance
-sci_inv
+e1=~eitem8+eitem10+eitem11+eitem12+eitem13+eitem16+eitem17+eitem20+
+   eitem21+eitem22+eitem23+eitem24+eitem25+eitem26+eitem28+eitem29+
+   eitem30+eitem31+eitem32+eitem33+eitem34+eitem36+eitem37
+e2=~eitem27+eitem35+eitem9+eitem18
+e3=~eitem38+eitem39+eitem40
+erawsc~e1+e2+e3
 @
-These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
-Others were excluded due to sample size and composition constraints.
-As a consequence of the $\chi$\textsuperscript{2} fit statistic scaling as a function of both sample size and the minimum function test statistic, traditional measurement invariance techniques of using changes in significance to reject a constraint fail here, as the sample size guarantees significant results.
-In response to this fact, we employ two different techniques: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
-For both, decreases in model fit of greater than .01 are considered significant.
-In addition, we include An (Akaike's) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps.
 
-\subsection{English and Mathematics}
 \subsubsection{Configural Invariance}
+<<>>=
+fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+@
 As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
 <<>>=
-fitMeasures(MI.model$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
 @
 
-\subsubsection{Metric Invariance}
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+<<>>=
+fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+\subsection{Mathematics}
+While the EFA presented a good model fit for a five factor model, when the SEM model imposed upon the data, it failed to converge due to underspecification.
+As such, factor 5 was removed, because it consisted of only four low loading items.
+<<>>=
+m1=~mitem4+mitem5+mitem8+mitem11+mitem12+mitem14+mitem25+mitem32+mitem35+mitem37+mitem40+mitem42
+m2=~mitem6+mitem10+mitem17+mitem24+mitem39
+m3=~mitem22+mitem28+mitem34+mitem38+mitem41
+m4=~mitem18+mitem20+mitem21+mitem31+mitem42
+mrawsc~m1+m2+m3+m4
+@
+
+\subsubsection{Configural Invariance}
+<<>>=
+fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
 <<>>=
-fitMeasures(MI.model$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
 @
 
-The second test demonstrates a failure of metric/weak invariance when comparing either the $\Delta$ TLI or $\Delta$ CFI.
-Commonly, metric invariance is used to demonstrate that for all items, one unit change in the item scores is scaled to an equal unit change in the factor score across groups.
-This implies that the variances on each measure are the same for all items.
-As a consequence of the failure of metric invarianc to hold, all subsequent equivalences are rejected as well.
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+<<>>=
+fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+@
 
@@
 In addition, each subsequent test produced similar results, indicating that across all four primary ethnicities, the test operated with equivalent measurement.
+Biology as presented an interpretable factor structure, which was represented with the following mathematical structure.
+<<>>=
+s1=~sitem1+sitem2+sitem13+sitem14+sitem15+sitem22+sitem27+sitem30+sitem33+sitem34+sitem37+sitem40+sitem41
+s2=~sitem3+sitem5+sitem18+sitem19+sitem28+sitem31+sitem35+sitem43+sitem45
+s3=~sitem8+sitem10+sitem11+sitem12+sitem23+sitem24+sitem25+sitem23
+s4=~sitem4+sitem23+sitem29+sitem32+sitem44
+srawsc~s1+s2+s3+s4
+@
+
+\subsubsection{Configural Invariance}
+<<>>=
+fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
 <<>>=
-fitMeasures(MI.model2$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
-fitMeasures(MI.model2$fit.means, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+<<>>=
+fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
 @
@@
 \subsection{Questionnaire}
+\subsubsection{Multivariate Regression Tree}
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters of the response variables.
 
-<<>>=
+<<fig=T>>=
 require(mvpart)
-fit1<-mvpart(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,data=d4)
-summary(manova(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
-summary(lm(cbind(d4$mrawsc,d4$erawsc,d4$srawsc)~.,d4))
+mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4,rsq=T)
 @
+
+These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
+The failure for the function to further discretize the high achievers indicates that the individual differences wihin accounted for no additional information.
+For lower aspiring students, Ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
+The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.
+Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15
+\subsubsection{Multivariate Regression}
+A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
 

### 2014-03-20 paper.tex
fixed final dataframes. All EFA and CFA code should work. text is incomplete atm. still need to find out how to plot means of each tree node, if possible. also, R code does not run in paper.rnw txt: Need to fix

@@
 \section{Results}
-\subsection{Theoretical structure of the test}
-Substantive theoretical knowledge of the number of latent knowledge domains that were intended to be tested was generated from the answer keys provided by Massachusetts' Department of Education.
-Specifically, there were 5 general topics within Biology:
-\begin{itemize}
-\item{Anatomy \& Physiology}
-\item{Biochemistry and Cell Biology}
-\item{Ecology}
-\item{Evolution and Biodiversity}
-\item{Genetics}
-\end{itemize}
-Two topics within English:
-\begin{itemize}
-\item{Reading \& Literature}
-\item{Language}
-\end{itemize}
-and five topics within Mathematics:
-\begin{itemize}
-\item{Data Analysis  Statistics  and Probability}
-\item{Geometry}
-\item{Measurement}
-\item{Number Sense and Operations}
-\item{Patterns Relations and Algebra}
-\end{itemize}
 \subsection{Exploratory factor analyses}
-With the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains, and factor loadings calculated using maximum likelihood estimates calculated using the tetrachoric covariance matrices for each subset of items.
-The psych package was utilised both to calculate the factor structure.
+Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
+The psych package was used to calculate the factor structure.
 Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
 \subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the exam, we chose to run two separate tests for measurement invariance by running Biology separately.
-Using the results of the EFA, we specified the following two models:
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
 \begin{knitrout}
@@
 \begin{alltt}
-\hlstd{invariance}
+\hlkwd{fa.diagram}\hlstd{(feng)}
 \end{alltt}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'invariance' not found}}\begin{alltt}
-\hlstd{sci_inv}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'sci\_inv' not found}}\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
 \end{knitrout}
-
-These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
-Others were excluded due to sample size and composition constraints.
-As a consequence of the $\chi$\textsuperscript{2} fit statistic scaling as a function of both sample size and the minimum function test statistic, traditional measurement invariance techniques of using changes in significance to reject a constraint fail here, as the sample size guarantees significant results.
-In response to this fact, we employ two different techniques: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
-For both, decreases in model fit of greater than .01 are considered significant.
-In addition, we include An (Akaike's) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps.
-
-\subsection{English and Mathematics}
-\subsubsection{Configural Invariance}
-As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
 
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fa.diagram}\hlstd{(fmat)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fa.diagram}\hlstd{(fsci1)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
+\end{knitrout}
+
+These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+Other ethnicities were excluded because of their relative rarity within the sample.
+As a consequence of the $\chi$\textsuperscript{2} fit statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0 \end{equation} scales as a function of both sample size and the minimum function test statistic $F_0$, traditional measurement invariance, which uses changes in significance to reject a constraint, fails here, as the sample size guarantees significant results.
+In response to this fact, we employ three different fitness criterion: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
+For all three, decreases in model fit of greater than .01 are considered significant reductions in model fit.
+In addition, we include An (Akaike) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps in comparison to the theoretical true model.
+
+\subsection{English}
+The English section constructed using the following system of latent and test items, as predictors of the unstandardized English score, as given below.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{e1} \hlkwb{=} \hlopt{~}\hlstd{eitem8} \hlopt{+} \hlstd{eitem10} \hlopt{+} \hlstd{eitem11} \hlopt{+} \hlstd{eitem12} \hlopt{+} \hlstd{eitem13} \hlopt{+} \hlstd{eitem16} \hlopt{+} \hlstd{eitem17} \hlopt{+} \hlstd{eitem20} \hlopt{+}
+    \hlstd{eitem21} \hlopt{+} \hlstd{eitem22} \hlopt{+} \hlstd{eitem23} \hlopt{+} \hlstd{eitem24} \hlopt{+} \hlstd{eitem25} \hlopt{+} \hlstd{eitem26} \hlopt{+} \hlstd{eitem28} \hlopt{+} \hlstd{eitem29} \hlopt{+}
+    \hlstd{eitem30} \hlopt{+} \hlstd{eitem31} \hlopt{+} \hlstd{eitem32} \hlopt{+} \hlstd{eitem33} \hlopt{+} \hlstd{eitem34} \hlopt{+} \hlstd{eitem36} \hlopt{+} \hlstd{eitem37}
+\hlstd{e2} \hlkwb{=} \hlopt{~}\hlstd{eitem27} \hlopt{+} \hlstd{eitem35} \hlopt{+} \hlstd{eitem9} \hlopt{+} \hlstd{eitem18}
+\hlstd{e3} \hlkwb{=} \hlopt{~}\hlstd{eitem38} \hlopt{+} \hlstd{eitem39} \hlopt{+} \hlstd{eitem40}
+\hlstd{erawsc} \hlopt{~} \hlstd{e1} \hlopt{+} \hlstd{e2} \hlopt{+} \hlstd{e3}
+\end{alltt}
+\begin{verbatim}
+## erawsc ~ e1 + e2 + e3
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\subsubsection{Configural Invariance}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 
-\subsubsection{Metric Invariance}
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 
-The second test demonstrates a failure of metric/weak invariance when comparing either the $\Delta$ TLI or $\Delta$ CFI.
-Commonly, metric invariance is used to demonstrate that for all items, one unit change in the item scores is scaled to an equal unit change in the factor score across groups.
-This implies that the variances on each measure are the same for all items.
-As a consequence of the failure of metric invarianc to hold, all subsequent equivalences are rejected as well.
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+\subsection{Mathematics}
+While the EFA presented a good model fit for a five factor model, when the SEM model imposed upon the data, it failed to converge due to underspecification.
+As such, factor 5 was removed, because it consisted of only four low loading items.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{m1} \hlkwb{=} \hlopt{~}\hlstd{mitem4} \hlopt{+} \hlstd{mitem5} \hlopt{+} \hlstd{mitem8} \hlopt{+} \hlstd{mitem11} \hlopt{+} \hlstd{mitem12} \hlopt{+} \hlstd{mitem14} \hlopt{+} \hlstd{mitem25} \hlopt{+} \hlstd{mitem32} \hlopt{+}
+    \hlstd{mitem35} \hlopt{+} \hlstd{mitem37} \hlopt{+} \hlstd{mitem40} \hlopt{+} \hlstd{mitem42}
+\hlstd{m2} \hlkwb{=} \hlopt{~}\hlstd{mitem6} \hlopt{+} \hlstd{mitem10} \hlopt{+} \hlstd{mitem17} \hlopt{+} \hlstd{mitem24} \hlopt{+} \hlstd{mitem39}
+\hlstd{m3} \hlkwb{=} \hlopt{~}\hlstd{mitem22} \hlopt{+} \hlstd{mitem28} \hlopt{+} \hlstd{mitem34} \hlopt{+} \hlstd{mitem38} \hlopt{+} \hlstd{mitem41}
+\hlstd{m4} \hlkwb{=} \hlopt{~}\hlstd{mitem18} \hlopt{+} \hlstd{mitem20} \hlopt{+} \hlstd{mitem21} \hlopt{+} \hlstd{mitem31} \hlopt{+} \hlstd{mitem42}
+\hlstd{m5} \hlkwb{=} \hlopt{~}\hlstd{mitem2} \hlopt{+} \hlstd{mitem3} \hlopt{+} \hlstd{mitem9} \hlopt{+} \hlstd{mitem18} \hlopt{+} \hlstd{mitem19}
+\hlstd{mrawsc} \hlopt{~} \hlstd{m1} \hlopt{+} \hlstd{m2} \hlopt{+} \hlstd{m3} \hlopt{+} \hlstd{m4} \hlopt{+} \hlstd{m5}
+\end{alltt}
+\begin{verbatim}
+## mrawsc ~ m1 + m2 + m3 + m4 + m5
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\subsubsection{Configural Invariance}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
 
@@
 In addition, each subsequent test produced similar results, indicating that across all four primary ethnicities, the test operated with equivalent measurement.
+Biology as presented an interpretable factor structure, which was represented with the following mathematical structure.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{s1} \hlkwb{=} \hlopt{~}\hlstd{sitem1} \hlopt{+} \hlstd{sitem2} \hlopt{+} \hlstd{sitem13} \hlopt{+} \hlstd{sitem14} \hlopt{+} \hlstd{sitem15} \hlopt{+} \hlstd{sitem22} \hlopt{+} \hlstd{sitem27} \hlopt{+} \hlstd{sitem30} \hlopt{+}
+    \hlstd{sitem33} \hlopt{+} \hlstd{sitem34} \hlopt{+} \hlstd{sitem37} \hlopt{+} \hlstd{sitem40} \hlopt{+} \hlstd{sitem41}
+\hlstd{s2} \hlkwb{=} \hlopt{~}\hlstd{sitem3} \hlopt{+} \hlstd{sitem5} \hlopt{+} \hlstd{sitem18} \hlopt{+} \hlstd{sitem19} \hlopt{+} \hlstd{sitem28} \hlopt{+} \hlstd{sitem31} \hlopt{+} \hlstd{sitem35} \hlopt{+} \hlstd{sitem43} \hlopt{+}
+    \hlstd{sitem45}
+\hlstd{s3} \hlkwb{=} \hlopt{~}\hlstd{sitem8} \hlopt{+} \hlstd{sitem10} \hlopt{+} \hlstd{sitem11} \hlopt{+} \hlstd{sitem12} \hlopt{+} \hlstd{sitem23} \hlopt{+} \hlstd{sitem24} \hlopt{+} \hlstd{sitem25} \hlopt{+} \hlstd{sitem23}
+\hlstd{s4} \hlkwb{=} \hlopt{~}\hlstd{sitem4} \hlopt{+} \hlstd{sitem23} \hlopt{+} \hlstd{sitem29} \hlopt{+} \hlstd{sitem32} \hlopt{+} \hlstd{sitem44}
+\hlstd{srawsc} \hlopt{~} \hlstd{s1} \hlopt{+} \hlstd{s2} \hlopt{+} \hlstd{s3} \hlopt{+} \hlstd{s4}
+\end{alltt}
+\begin{verbatim}
+## srawsc ~ s1 + s2 + s3 + s4
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\subsubsection{Configural Invariance}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+\subsubsection{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.model2}\hlopt{$}\hlstd{fit.means,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\subsubsection{Strong Invariance}
+Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
+, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
+As a consequence, strict invariance is not tested, because by definition it cannot hold.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 \subsection{Questionnaire}
-Preliminary results were conducted using a multivariate regression tree, which searches for both linear and non-linear relationships between different classifications of the three raw scores and the questionnaire items, ethncity, and gender.
-However, the data failed to split at all, meaning that there is little information that relates between the questionnaire, demographics, and individuals' scores.
-Likewise mutlivariate regression results included demonstrate the no meaningful results were identified, with the few significant results conceptually meaningless.
+\subsubsection{Multivariate Regression Tree}
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters of the response variables.
+
 \begin{knitrout}
@@
 {\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: mvpart}}\begin{alltt}
-\hlstd{fit1} \hlkwb{<-} \hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(d}\hlopt{$}\hlstd{mrawsc, d}\hlopt{$}\hlstd{erawsc, d}\hlopt{$}\hlstd{srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'd4' not found}}\begin{alltt}
-\hlkwd{summary}\hlstd{(}\hlkwd{manova}\hlstd{(}\hlkwd{cbind}\hlstd{(d}\hlopt{$}\hlstd{mrawsc, d}\hlopt{$}\hlstd{erawsc, d}\hlopt{$}\hlstd{srawsc)} \hlopt{~} \hlstd{., d4))}
+\hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(mrawsc, erawsc, srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4,} \hlkwc{rsq} \hlstd{= T)}
 \end{alltt}
@@
 \end{knitrout}
+
+These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
+The failure for the function to further discretize the high achievers indicates that the individual differences wihin accounted for no additional information.
+For lower aspiring students, Ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
+The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.
+Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15
+\subsubsection{Multivariate Regression}
+A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
 

### 2014-03-20 EDA work.R

@@ -36,9 +36,10 @@ sci1<-d[d$scitry==1,c(95:140,170)]
 require(gdata)
 sci1<-as.data.frame(lapply(sci1,drop.levels))
 scix<-sci1[,c(13,24,33,45,46,47)]
+sci1$sitem12<-sci1$sitem32<-NULL
 sci1<-as.data.frame(cbind(codive(sci1[sapply(sci1,is.factor)]),sci1$id))
-sci1$id<-sci1$V43
-sci1$V43<-NULL
+sci1$id<-sci1$V41
+sci1$V41<-NULL
 
 scix[scix==' ']<-NA
 scix<-as.data.frame(lapply(scix,as.integer))
@@ -50,7 +51,6 @@ scixx<-sci2[,c(12,24,34,45,46,47)]
 sci2<-as.data.frame(cbind(codive(sci2[sapply(sci2,is.factor)]),sci2$id))
 sci2$id<-sci2$V43
 sci2$V43<-NULL
-
 scixx[scixx==' ']<-NA
 scixx<-as.data.frame(lapply(scixx,as.integer))
 sci2<-merge(sci2,scixx,by='id')
@@ -78,36 +78,38 @@ scixxxx<-as.data.frame(lapply(scixxxx,as.integer))
 sci4<-merge(sci4,scixxxx,by='id')
 
 require(psych)
-feng<-fa.poly(eng,nfactors=3,n.iter=1000,rotate='oblimin',fm='ml',oblique.scores=T)
+feng<-fa.poly(eng[,c(2:41)],nfactors=3,n.iter=10,rotate='oblimin',fm='ml',oblique.scores=T)
 # fengh<-omega(poe,nfactors=3,n.iter=1000,rotate='oblimin',fm='ml',sl=F)
 fa.diagram(feng)
 fmat<-fa.poly(mat,nfactors=5,n.iter=1000,rotate='oblimin',fm='ml',oblique.scores=T)
 
-fsci1<-fa.poly(sci1,nfactors=4,n.iter=1,rotate='oblimin',fm='ml',oblique.scores=T)
+fsci1<-fa.poly(sci1,nfactors=4,n.iter=100,rotate='oblimin',fm='ml',oblique.scores=T)
 
 ##Below gives us two of the three domains to work with.
 two<-merge(eng,mat,by='id')
[…]

### 2014-03-20 paper-knitr.Rnw
updates et cet.

--- 2014-03-20 paper.Rnw
@@
 <<fig=T>>=
-fa.diagram(feng)
+setwd('~/Documents/workspace/DataCompete/upstat-competition')
+fa.diagram(feng,main='English Oblimin')
 @
@@
 <<fig=T>>=
-fa.diagram(fmat)
+fa.diagram(fmat,main='Mathematics Oblimin')
 @
@@
 <<fig=T>>=
-fa.diagram(fsci1)
+fa.diagram(fsci1,main='Biology Oblimin')
 @
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
 <<>>=
@@
 
+\subsubsection{Strict Invariance}
+Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
+<<>>=
+fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
+@
 \subsection{Mathematics}
@@
 <<>>=
-fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.maths$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
 @
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
 
 <<>>=
-fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.maths$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
 @
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance does not substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000, the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=167.652.
+
 <<>>=
-fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
+\subsubsection{Strict Invariance}
+Strong Invariance imposes a significant constraint upon model fit: $\Delta$ CFI = .017, $\Delta$ RMSEA = .005, the $\Delta$ TLI = .014 and $\Delta$$\chi^2$=998.348.
+
+<<>>=
+fitMeasures(MI.maths$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
 @
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance does fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875.
 <<>>=
@@
 
+\subsubsection{Strict Invariance}
+Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
+<<>>=
+fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
+@
+
 \subsection{Questionnaire}
 \subsubsection{Multivariate Regression Tree}
-Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters of the response variables.
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
+The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
+We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate nonresponse as a partitioning criterion.
 
@@
 These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
-The failure for the function to further discretize the high achievers indicates that the individual differences wihin accounted for no additional information.
-For lower aspiring students, Ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
+The failure for the function to further discretize the low achievers indicates that the individual differences wihin accounted for no additional information.
+For high aspiring students, ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
 The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.

### 2014-03-20 EDA work.R

@@ -111,14 +111,6 @@ require(gdata)
 d3$eth<-drop.levels(d3$eth,reorder=T)
 d2$eth<-drop.levels(d2$eth,reorder=T)
 
-invariance="f1=~eitem8+eitem13+eitem22+eitem26+eitem28+eitem30+eitem31
-            f2=~eitem38+eitem39+eitem40
-            f4=~mitem40+mitem42+mitem35+mitem12+mitem11+mitem5
-            f5=~mitem6+mitem10+mitem17+mitem39
-            f6=~mitem22+mitem28+mitem34+mitem38+mitem31
-            f7=~mitem18+mitem20+mitem21+mitem31+mitem41
-            erawsc~f1+f2
-            mrawsc~f4+f5+f6+f7"
 engl<-'
 e1=~eitem8+eitem10+eitem11+eitem12+eitem13+eitem16+eitem17+eitem20+eitem21+eitem22+eitem23+eitem24+eitem25+eitem26+eitem28+eitem29+eitem30+eitem31+eitem32+eitem33+eitem34+eitem36+eitem37
 e2=~eitem27+eitem35+eitem9+eitem18
@@ -136,47 +128,49 @@ s4=~sitem4+sitem23+sitem29+sitem32+sitem44
 srawsc~s1+s2+s3+s4'
   
   
-MI.model<-measurementInvariance(invariance,d2,group='eth',missing='fiml',bootstrap=1000)
-MI.all<-measurementInvariance(all,d3,group='eth',missing='fiml',bootstrap=100)
+MI.eng<-measurementInvariance(engl,d3,group='eth',missing='fiml',bootstrap=150,strict=T)
+MI.maths<-measurementInvariance(maths,d3,group='eth',missing='fiml',bootstrap=150,strict=T)
+MI.scien<-measurementInvariance(scien,d3,group='eth',missing='fiml',bootstrap=150,strict=T)
+
+fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
+fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
[…]

### 2014-03-20 paper.tex
updates et cet.

@@
 \begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(feng)}
+\hlkwd{setwd}\hlstd{(}\hlstr{"~/Documents/workspace/DataCompete/upstat-competition"}\hlstd{)}
+\hlkwd{fa.diagram}\hlstd{(feng,} \hlkwc{main} \hlstd{=} \hlstr{"English Oblimin"}\hlstd{)}
 \end{alltt}
@@
 \begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fmat)}
+\hlkwd{fa.diagram}\hlstd{(fmat,} \hlkwc{main} \hlstd{=} \hlstr{"Mathematics Oblimin"}\hlstd{)}
 \end{alltt}
@@
 \begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fsci1)}
+\hlkwd{fa.diagram}\hlstd{(fsci1,} \hlkwc{main} \hlstd{=} \hlstr{"Biology Oblimin"}\hlstd{)}
 \end{alltt}
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
 \begin{knitrout}
@@
 \hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+\subsubsection{Strict Invariance}
+Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.residuals,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 \hlstd{m4} \hlkwb{=} \hlopt{~}\hlstd{mitem18} \hlopt{+} \hlstd{mitem20} \hlopt{+} \hlstd{mitem21} \hlopt{+} \hlstd{mitem31} \hlopt{+} \hlstd{mitem42}
-\hlstd{m5} \hlkwb{=} \hlopt{~}\hlstd{mitem2} \hlopt{+} \hlstd{mitem3} \hlopt{+} \hlstd{mitem9} \hlopt{+} \hlstd{mitem18} \hlopt{+} \hlstd{mitem19}
-\hlstd{mrawsc} \hlopt{~} \hlstd{m1} \hlopt{+} \hlstd{m2} \hlopt{+} \hlstd{m3} \hlopt{+} \hlstd{m4} \hlopt{+} \hlstd{m5}
+\hlstd{mrawsc} \hlopt{~} \hlstd{m1} \hlopt{+} \hlstd{m2} \hlopt{+} \hlstd{m3} \hlopt{+} \hlstd{m4}
 \end{alltt}
 \begin{verbatim}
-## mrawsc ~ m1 + m2 + m3 + m4 + m5
+## mrawsc ~ m1 + m2 + m3 + m4
 \end{verbatim}
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
 
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance does not substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000, the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=167.652.
+
 \begin{knitrout}
@@
 \begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
+\subsubsection{Strict Invariance}
+Strong Invariance imposes a significant constraint upon model fit: $\Delta$ CFI = .017, $\Delta$ RMSEA = .005, the $\Delta$ TLI = .014 and $\Delta$$\chi^2$=998.348.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.residuals,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
 \end{alltt}
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA =.
-, and the $\Delta$ TLI = and $\Delta$$\chi^2$=187.875.
-As a consequence, strict invariance is not tested, because by definition it cannot hold.
+Strong Invariance does fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875.
 \begin{knitrout}
@@
 
+\subsubsection{Strict Invariance}
+Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.residuals,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\end{knitrout}
+
 \subsection{Questionnaire}
 \subsubsection{Multivariate Regression Tree}
-Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters of the response variables.
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
+The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
 

### 2014-03-21 paper.Rnw
updates et cet.
update with figures

@@
 The data also contains basic demographic information for each student comprised of race/ethnicity, gender, and an academic engagement questionnaire.
-Each test consists of multiple choice items and open ended type questions that are graded on a holistic scale, implemented as a 5 point integer scale.
+Each test consists of multiple choice items and open-ended questions that are graded on a holistic scale, implemented as a 5 point integer scale.
 
@@
 Tests whether in addition to the same factor structure, the same items load equivalently onto the same same structure.
-\item{Strong invariance}
+\item{Strong invariance:}
 Tests that the intercepts, item loadings, and factor structure are equivalent across groups.
@@
 These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
-In this paper, we examine these factors using a series of R packages found cited at the end of the paper \cite{psych,lavaan,semTools}.
+In this paper, we examine these factors using a group of R packages cited at the end of the paper \cite{psych,lavaan,semTools}.
 Exploratory bootstrapped oblimin factor analyses were conducted upon each section to establish the number of domains and high loading items within each of the three tests.
@@
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
 However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
@@
 As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
-<<fig=T>>=
-setwd('~/Documents/workspace/DataCompete/upstat-competition')
-fa.diagram(feng,main='English Oblimin')
-@
+
+\includegraphics{/home/landon/Documents/workspace/DataCompete/upstat-competition/figure/English.jpeg}
 
@@
 \subsubsection{Multivariate Regression Tree}
+
 Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
@@
 require(mvpart)
-mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4,rsq=T)
+mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4)
 @
@@
 \subsubsection{Multivariate Regression}
-A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
+A parametric regression model was also conducted exploring only main effects between the predictors and response.
 
@@
 As indicated in the results section, the test fails to perform equivalently across the four groups, which is the primary point of order in test design and advocacy.
-This lends substantive credence to the argument that there are fundamental flaws within the test.
-However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
-In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between the general education practices taught, which is likely a function of SES, which disproportionately affects minorities over Whites.
-This subtle distinction has the immediate consequence of arguing that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
-
-Little more can be said about the questionnaire beyond the fact that the mode for each item is the most efficient predictor of individual student responses.
-As such, we can only conclude that the information being measured by these items are independent of the exams' item functioning and student performance.
+Moreover, strict invariance has a specific import upon the function of psychometric and education measurements, testing whether the error variances --the portions of item variation not attributable to the common factors.
+Conceptually, strict invariance validates the assumption that error variances are different across groups; if this fails, there are either different variables operating on the measures across groups or the same set of variables operates differently across groups \cite{deshon}.
+This fundamentally demonstrates the importance of ensuring that all four mainstays of measurement invariance be substantiated, especially when one considers the long standing view that residuals convey meaning of unmeasured latent constructs' effects, moderated by group membership in this case \cite{cronbach}.
+This lends substantive credence to the argument that there are fundamental flaws within the test, specifically with reflection towards students' socioeconomic status.
 

### 2014-03-21 paper.tex
update with figures

@@
 As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{setwd}\hlstd{(}\hlstr{"~/Documents/workspace/DataCompete/upstat-competition"}\hlstd{)}
-\hlkwd{fa.diagram}\hlstd{(feng,} \hlkwc{main} \hlstd{=} \hlstr{"English Oblimin"}\hlstd{)}
-\end{alltt}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
-\end{knitrout}
+\includegraphics{/home/landon/Documents/workspace/DataCompete/upstat-competition/figure/English.jpeg}
 
@@
 \subsubsection{Multivariate Regression Tree}
+
 Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
+We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate nonresponse as a partitioning criterion.
 
@@
 {\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: mvpart}}\begin{alltt}
-\hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(mrawsc, erawsc, srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4,} \hlkwc{rsq} \hlstd{= T)}
+\hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(mrawsc, erawsc, srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4)}
 \end{alltt}

### 2014-03-24 paper-knitr.Rnw
edits and fixes

@@
 
-<<setup, include=T, cache=FALSE, echo=FALSE>>=
+<<setup, include=F, cache=FALSE, echo=FALSE>>=
+load('/home/landon/Documents/workspace/DataCompete/upstat-competition/.RData')
+require(psych)
+require(semTools)
+require(mvpart)
 Sys.setenv(TEXINPUTS=getwd(), 
@@
 
-The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
+The Massachusetts Comprehensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
 We examine a subset sample of 10\textsuperscript{th} grade students' results from the 2009 Spring MCAS test.
 The data also contains basic demographic information for each student comprised of race/ethnicity, gender, and an academic engagement questionnaire.
-Each test consists of multiple choice items and open ended type questions that are graded on a holistic scale, implemented as a 5 point integer scale.
+Each test consists of multiple choice items and open-ended questions that are graded on a holistic scale, implemented as a 5 point integer scale.
 
@@
 The former, comprised mainly of teachers and educational policy analysts, claim that the test is unfairly biased against students as a function of demographics, while citing the significant impediment that poor performance upon the tests has upon both immediate and long-term outcomes.
-Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
+Theoretical perspectives on test design are the primary domain of psychometric, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealistic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
 
 \subsection{Theoretical structure of the test}
-Our paper begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+Our paper begins with an analysis of both the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 \section{Methods}
 \subsection{Subjects}
-Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusett's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} graders.
+Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusetts's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} graders.
 Provided demographics present as follows 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female, and there are no significant interactions between gender and ethnicity.
 
-Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relaionship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
+Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relationship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
 \subsection{Procedure}
-A critical assumption within standardised testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
+A critical assumption within standardized testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
 This property is a series of mathematical constraints applied to latent variables that are assumed to be non-significantly different between groups.
-Latent variables are unobservable constructs that underly the observed measures; one typically estiamtes them by computing standard principal axis factor analysis.
+Latent variables are unobservable constructs that underlie the observed measures; one typically estimates them by computing standard principal axis factor analysis.
 The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample, by imposing restrictions in a sequential four phase procedure:
@@
 Tests whether in addition to the same factor structure, the same items load equivalently onto the same same structure.
-\item{Strong invariance}
+\item{Strong invariance:}
 Tests that the intercepts, item loadings, and factor structure are equivalent across groups.
 \item{Strict invariance:}
-Tests the assumption that in addition to the preceeding steps, the residual variances are equivalent across groups.
+Tests the assumption that in addition to the preceding steps, the residual variances are equivalent across groups.
 \end{enumerate}
 These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
-In this paper, we examine these factors using a series of R packages found cited at the end of the paper \cite{psych,lavaan,semTools}.
+In this paper, we examine these factors using a group of R packages cited at the end of the paper \cite{psych,lavaan,semTools}.
 Exploratory bootstrapped oblimin factor analyses were conducted upon each section to establish the number of domains and high loading items within each of the three tests.
@@
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
-However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
-Even so, for the MG-CFA, we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation, in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+However, within the structure of the exam, missing data was unsurprisingly a rare occurrence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
+Even so, for the MG-CFA, we attempted to account for this by utilizing the Full Information Maximum Likelihood (FIML) estimation, in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
 However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques, producing potentially biased results under the assumption that items are missing completely at random.
@@
 \subsection{Exploratory factor analyses}
-Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
+Given the preceding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
-\subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
-<<fig=T>>=
-setwd('~/Documents/workspace/DataCompete/upstat-competition')
+<<fig=T,echo=F>>=
 fa.diagram(feng,main='English Oblimin')
-@
-
-<<fig=T>>=
 fa.diagram(fmat,main='Mathematics Oblimin')
-@
-
-<<fig=T>>=
 fa.diagram(fsci1,main='Biology Oblimin')
 @
+\subsection{MG-CFA}
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run three separate tests for measurement invariance:
 
@@
 The English section constructed using the following system of latent and test items, as predictors of the unstandardized English score, as given below.
-<<>>=
+<<echo=F>>=
 e1=~eitem8+eitem10+eitem11+eitem12+eitem13+eitem16+eitem17+eitem20+
@@
 \subsubsection{Configural Invariance}
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 As such, factor 5 was removed, because it consisted of only four low loading items.
-<<>>=
+<<echo=F>>=
 m1=~mitem4+mitem5+mitem8+mitem11+mitem12+mitem14+mitem25+mitem32+mitem35+mitem37+mitem40+mitem42
@@
 \subsubsection{Configural Invariance}
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.maths$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.maths$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<>>=
-fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+<<echo=F>>=
+round(fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic')),4)
 @
@@
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.maths$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Biology as presented an interpretable factor structure, which was represented with the following mathematical structure.
-<<>>=
+<<echo=F>>=
 s1=~sitem1+sitem2+sitem13+sitem14+sitem15+sitem22+sitem27+sitem30+sitem33+sitem34+sitem37+sitem40+sitem41
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strong Invariance does fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875.
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Multivariate Regression Tree}
+
 Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
-We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate nonresponse as a partitioning criterion.
+We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate non-response as a partitioning criterion.
 
-<<fig=T>>=
-require(mvpart)
-mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4,rsq=T)
+<<echo=F,fig=T>>=
+mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4)
 @
@@
 These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
-The failure for the function to further discretize the low achievers indicates that the individual differences wihin accounted for no additional information.
+The failure for the function to further discretize the low achievers indicates that the individual differences within accounted for no additional information.
 For high aspiring students, ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
@@
 As indicated in the results section, the test fails to perform equivalently across the four groups, which is the primary point of order in test design and advocacy.
-This lends substantive credence to the argument that there are fundamental flaws within the test.
-However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
-In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between the general education practices taught, which is likely a function of SES, which disproportionately affects minorities over Whites.
-This subtle distinction has the immediate consequence of arguing that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
+Moreover, strict invariance has a specific import upon the function of psychometric and education measurements, testing whether the error variances --the portions of item variation not attributable to the common factors.
+Conceptually, strict invariance validates the assumption that error variances are different across groups; if this fails, there are either different variables operating on the measures across groups or the same set of variables operates differently across groups \cite{deshon}.
+This fundamentally demonstrates the importance of ensuring that all four mainstays of measurement invariance be substantiated, especially when one considers the long standing view that residuals convey meaning of unmeasured latent constructs' effects, moderated by group membership in this case \cite{cronbach}.
+This lends substantive credence to the argument that there are fundamental flaws within the test, specifically with reflection towards students' socioeconomic status.
 
-Little more can be said about the questionnaire beyond the fact that the mode for each item is the most efficient predictor of individual student responses.
-As such, we can only conclude that the information being measured by these items are independent of the exams' item functioning and student performance.
+The questionnaire itself holds some important implications for students, namely that increased exposure quality education results in better test performance (c.f., question 14).
+The low value of predictive accuracy, found within both sets of results, indicates that these survey demographics are either largely irrelevant to the performance of a student, or that they are inaccurate measures of performance.
+The fact that survey non-response is a significant classifier of above average performing students would indicate that meaningful results are unlikely to be found.
 

### 2014-03-24 EDA work.R

@@ -123,7 +123,7 @@ m4=~mitem18+mitem20+mitem21+mitem31+mitem42
 mrawsc~m1+m2+m3+m4'
 scien<-'s1=~sitem1+sitem2+sitem13+sitem14+sitem15+sitem22+sitem27+sitem30+sitem33+sitem34+sitem37+sitem40+sitem41
 s2=~sitem3+sitem5+sitem18+sitem19+sitem28+sitem31+sitem35+sitem43+sitem45
-s3=~sitem8+sitem10+sitem11+sitem12+sitem23+sitem24+sitem25+sitem23
+s3=~sitem8+sitem10+sitem11+sitem12+sitem23+sitem24+sitem25
 s4=~sitem4+sitem23+sitem29+sitem32+sitem44
 srawsc~s1+s2+s3+s4'
   
@@ -151,7 +151,7 @@ fitMeasures(MI.scien$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
 fitMeasures(MI.scien$fit.means, c('chisq','tli', 'cfi','rmsea','aic'))
 
 
-d4<-d[,c(145:172,45,91,141,95)]
+d4<-d[,c(145:172,45,91,141,95,2)]
 require(mvpart)
 fit<-mvpart(cbind(mrawsc,erawsc,srawsc)~.,d4)
 
@@ -166,8 +166,9 @@ d5$race_off<-NULL
 
 require(mice)
 d6<-mice(d5,m=13,maxit=15,defaultMethod = c("norm","logreg","polyreg"))
-summary(manova(cbind(mrawsc,erawsc,srawsc)~.,d5))
-summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5))
+summary(aggr(d5[,c(1,7:31)]))
+summary(manova(cbind(mrawsc,erawsc,srawsc)~.,d5)) #this is on the data with missing 
+summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5)) #this is on the data with missing 
 
 summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~.)))
 summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~.)))

### 2014-03-24 paper-knitr.Rnw
Added abstract, fixed some typos, figure modifications

--- 2014-03-20 paper-knitr.Rnw
@@
 
-<<setup, include=T, cache=FALSE, echo=FALSE>>=
+<<setup, include=FALSE, cache=FALSE, echo=FALSE, results='hide'>>=
+library(lavaan)
+library(psych)
+library(mvpart)
+
+load("comp.RData")
+
 Sys.setenv(TEXINPUTS=getwd(), 
@@
 
+The Massachusetts Comprehensive Assessment System (MCAS) is a set of standardized tests administered to Massachusetts public school students.
+We performed a series of statistical analyses on a subset of 10\textsuperscript{th} grade student test results in English, Mathematics, and Biology tests.
+An important property of any testing procedure is that of measurement invariance: the test should measure the same underlying constructs, in the same manner, regardless of what groups (ethnicicy, gender, etc.) each testee belongs to.
+We employed Structural Equation Modeling (SEM) to test for measurement invariance across four ethnic groups (White, Black, Hispanic, and Asian) in the MCAS results.
+Our results indicate that the MCAS does not function similarly across the ethnic groups, a result which undermines the validity of the test.
+Additionally, we applied multivariate regression and multivariate regression trees to analyze results from a questionnaire given to testees.
+Our results indicate that students who indicate ambition to pursue a bachelor's degree have higher scores on each test and are significantly affected by higher quality educational programs.
+
 \section{Introduction}
@@
 
-The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
+The Massachusetts Comprehensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
 We examine a subset sample of 10\textsuperscript{th} grade students' results from the 2009 Spring MCAS test.
@@
 The former, comprised mainly of teachers and educational policy analysts, claim that the test is unfairly biased against students as a function of demographics, while citing the significant impediment that poor performance upon the tests has upon both immediate and long-term outcomes.
-Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
+\cite{Gaudet} Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
 
 \subsection{Theoretical structure of the test}
-Our paper begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+Our paper begins with an analysis of the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+
 \section{Methods}
@@
 Our sample is comprised of 10,515 10\textsuperscript{th} grade students enrolled in Massachusett's public education system, and is assumed to be drawn as an unbiased sample of the total universe of 10\textsuperscript{th} graders.
-Provided demographics present as follows 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female, and there are no significant interactions between gender and ethnicity.
+Provided demographics present as follows: 4.48\% Asian, 2.97\% Black, 3.63\% Hispanic/Latino, 1.8\% multiracial, 0.17\% Native American, .12\% Pacific Islander, and 86.82\% White; 50.32\% was female, and there are no significant interactions between gender and ethnicity.
 
 Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relaionship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
+
 \subsection{Procedure}
@@
 This property is a series of mathematical constraints applied to latent variables that are assumed to be non-significantly different between groups.
-Latent variables are unobservable constructs that underly the observed measures; one typically estiamtes them by computing standard principal axis factor analysis.
-The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample, by imposing restrictions in a sequential four phase procedure:
+Latent variables are unobservable constructs that underly the observed measures; one typically estimates them by computing standard principal axis factor analysis.
+The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample.
+This is shown by imposing restrictions in a sequential four phase procedure:
 \begin{enumerate}
@@
 \item{Weak invariance:}
-Tests whether in addition to the same factor structure, the same items load equivalently onto the same same structure.
+Tests whether in addition to the same factor structure, the same items load equivalently onto the same structure.
 \item{Strong invariance}
@@
 These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
-In this paper, we examine these factors using a series of R packages found cited at the end of the paper \cite{psych,lavaan,semTools}.
-Exploratory bootstrapped oblimin factor analyses were conducted upon each section to establish the number of domains and high loading items within each of the three tests.
+In this paper, we examine these factors using a series of R packages \cite{psych,lavaan,semTools}.
+Exploratory bootstrapped oblimin factor analyses were conducted upon each test to establish the number of domains and high loading items within each of the three tests.
 As a consequence of the extreme ethnicity imbalance within three of the four science exams, only Biology was tested using MG-CFA.
 
-Furthermore, we investigated the information found within the questionnaire, relating them to the students' scores on the three sections using both multivariate parametric regressions, and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
+Furthermore, we investigated the information found within the questionnaire by relating them to the students' scores on the three sections.
+We used both multivariate parametric regressions and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
 However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
-Even so, for the MG-CFA, we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation, in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
-However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques, producing potentially biased results under the assumption that items are missing completely at random.
-As such, we processed the original responses using the mice package\cite{mice}, using 25 imputed data frames with 30 iterations in each dataframe.
-Regression trees are relatively robust to missing data, because all information is not processed simultaneously, but sequentially in order of variable importance, and as such we did not take any additional steps when conducting that analysis.
+Even so, for the MG-CFA we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
+As such, we processed the original responses using the mice package\cite{mice} using 25 imputed data frames with 30 iterations in each dataframe.
+Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.
+As such we did not take any additional steps when conducting that analysis.
 \section{Results}
 \subsection{Exploratory factor analyses}
-Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
+Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
+Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
+Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix A.
+
 \subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
-<<fig=T>>=
-setwd('~/Documents/workspace/DataCompete/upstat-competition')
-fa.diagram(feng,main='English Oblimin')
-@
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239).
+Diagrams representing the EFA results can be found in Appendix A.
 
-<<fig=T>>=
-fa.diagram(fmat,main='Mathematics Oblimin')
-@
+ The latent variable models for each test were specified using a FIML procedure with
+bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+Other ethnicities were excluded because of their relative rarity within
+the sample. As a consequence of the $\chi$\textsuperscript{2} fit
+statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0
+\end{equation} scales as a function of both sample size and the minimum
+function test statistic $F_0$, traditional measurement invariance, which
+uses changes in significance to reject a constraint, fails here, as the
+sample size guarantees significant results. In response to this fact, we
+employ three different fitness criterion: the $\Delta$ Comparative Fit
+Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of
+approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI)
+\cite{Chen}. For all three, decreases in model fit of greater than .01
+are considered significant reductions in model fit. In addition, we
+include An (Akaike) Information Criterion (AIC), to demonstrate whether
+model fit increases or decreases between sequential steps in comparison
+to the theoretical true model.
+\subsection{English}
+The English section was constructed using the system of latent and test items presented below as predictors of the unstandardized English score.
 
-<<fig=T>>=
-fa.diagram(fsci1,main='Biology Oblimin')
-@
+\vspace{0.25cm}
 
-These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
-Other ethnicities were excluded because of their relative rarity within the sample.
-As a consequence of the $\chi$\textsuperscript{2} fit statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0 \end{equation} scales as a function of both sample size and the minimum function test statistic $F_0$, traditional measurement invariance, which uses changes in significance to reject a constraint, fails here, as the sample size guarantees significant results.
-In response to this fact, we employ three different fitness criterion: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
-For all three, decreases in model fit of greater than .01 are considered significant reductions in model fit.
-In addition, we include An (Akaike) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps in comparison to the theoretical true model.
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 8, 10, 11, 12, 13, 16, 17, 20,  21, 22, 23, 24, 25, 26, 28, 29,  30,
+31, 32, 33, 34, 36, 37\tabularnewline
+\hline 
+2 & 27, 35, 9, 18\tabularnewline
+\hline 
+3 & 38, 39, 40\tabularnewline
+\hline
+\end{tabular}
 
-\subsection{English}
-The English section constructed using the following system of latent and test items, as predictors of the unstandardized English score, as given below.
-<<>>=
-e1=~eitem8+eitem10+eitem11+eitem12+eitem13+eitem16+eitem17+eitem20+
-   eitem21+eitem22+eitem23+eitem24+eitem25+eitem26+eitem28+eitem29+
-   eitem30+eitem31+eitem32+eitem33+eitem34+eitem36+eitem37
-e2=~eitem27+eitem35+eitem9+eitem18
-e3=~eitem38+eitem39+eitem40
-erawsc~e1+e2+e3
-@
+\bigskip
+
+All latent variable indicators are questions from the English section of the test (e.g. `8' represents `eitem8').
+
+\end{minipage}
 
 \subsubsection{Configural Invariance}
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI = 0.000.
 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
-<<>>=
+
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
-<<>>=
+
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsection{Mathematics}
-While the EFA presented a good model fit for a five factor model, when the SEM model imposed upon the data, it failed to converge due to underspecification.
-As such, factor 5 was removed, because it consisted of only four low loading items.
-<<>>=
-m1=~mitem4+mitem5+mitem8+mitem11+mitem12+mitem14+mitem25+mitem32+mitem35+mitem37+mitem40+mitem42
-m2=~mitem6+mitem10+mitem17+mitem24+mitem39
-m3=~mitem22+mitem28+mitem34+mitem38+mitem41
-m4=~mitem18+mitem20+mitem21+mitem31+mitem42
-mrawsc~m1+m2+m3+m4
-@
+While the EFA presented a good model fit for a five factor model, when the SEM model was imposed upon the data it failed to converge due to underspecification.
+As such, factor 5 was removed because it consisted of only four low loading items.
+The latent variable model for Mathematics is presented in the table below.
+
+\vspace{0.25cm}
+
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 4, 5, 8, 11, 12, 14, 25, 32, 35, 37, 40, 42\tabularnewline
+\hline 
+2 & 6, 10, 17, 24, 39\tabularnewline
+\hline 
+3 & 22, 28, 34, 38, 41\tabularnewline
+\hline 
+4 & 18, 20, 21, 31, 42\tabularnewline
+\hline 
+\end{tabular}
+
+\bigskip
+
+All latent variable indicators are questions from the Mathematics section of the test (e.g. `8' represents `mitem8').
+
+\end{minipage}
 
 \subsubsection{Configural Invariance}
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.maths$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
 @
-As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+The summary statistics presented below show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
 \subsubsection{Metric (Weak) Invariance}
@@
 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.maths$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.maths$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 In addition, each subsequent test produced similar results, indicating that across all four primary ethnicities, the test operated with equivalent measurement.
-Biology as presented an interpretable factor structure, which was represented with the following mathematical structure.
-<<>>=
-s1=~sitem1+sitem2+sitem13+sitem14+sitem15+sitem22+sitem27+sitem30+sitem33+sitem34+sitem37+sitem40+sitem41
-s2=~sitem3+sitem5+sitem18+sitem19+sitem28+sitem31+sitem35+sitem43+sitem45
-s3=~sitem8+sitem10+sitem11+sitem12+sitem23+sitem24+sitem25+sitem23
-s4=~sitem4+sitem23+sitem29+sitem32+sitem44
-srawsc~s1+s2+s3+s4
+Biology presented an interpretable factor structure, which was represented with the following mathematical structure.
+
+\vspace{0.25cm}
+
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 1, 2, 13, 14, 15, 22, 27, 30, 33, 34, 37, 40, 41\tabularnewline
+\hline 
+2 & 3, 5, 18, 19, 28, 31, 35, 43, 45\tabularnewline
+\hline 
+3 & 8, 10, 11, 12, 23, 24, 25\tabularnewline
+\hline 
+4 & 4, 23, 29, 32, 44\tabularnewline
+\hline 
+\end{tabular}
+
+\bigskip
+
+All latent variable indicators are questions from the Biology section of the test (e.g. `8' represents `sitem8').
+
+\end{minipage}
+
+\subsubsection{Configural Invariance}
+The analyses' summary statistics show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
+
+<<echo=FALSE>>=
+fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
 @
 
-\subsubsection{Configural Invariance}
-<<>>=
-fitMeasures(MI.eng$fit.configural, c('chisq','tli', 'cfi','rmsea','aic'))
-@
-As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
+While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875.
-<<>>=
+Strong Invariance fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875. 
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
-<<>>=
+<<echo=FALSE>>=
 fitMeasures(MI.eng$fit.residuals, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 \subsubsection{Multivariate Regression Tree}
-Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath} which is conceptually similar to running a multinomial regression upon the different clusters of the response variables.
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
@@
 
-<<fig=T>>=
+<<fig=T, echo=FALSE, fig.height=4>>=
 require(mvpart)
-mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4,rsq=T)
+mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4, legend=TRUE)
 @
@@
 These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
-The failure for the function to further discretize the low achievers indicates that the individual differences wihin accounted for no additional information.
-For high aspiring students, ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
+The failure for the function to further discretize the low achievers indicates that the individual differences within this group accounted for no additional information.
+For high aspiring students, ethnicity was found to be the second most important variable in discriminating test scores: Blacks, Hispanics and Native Americans performed similarly worse on the Mathematics and Biology sections compared to the other ethnicities.
 The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.
 Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15
+
 \subsubsection{Multivariate Regression}
@@
 \section{Discussion}
-As indicated in the results section, the test fails to perform equivalently across the four groups, which is the primary point of order in test design and advocacy.
-This lends substantive credence to the argument that there are fundamental flaws within the test.
+
+The primary point of order in test design is for a test to perform equivalently across groups of testees.
+As indicated in the results section, our analysis suggests that the test fails to perform equivalently across four ethnicity groups.
+This supplies quantitative evidence to the argument that there are fundamental flaws within the test.
 However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
-In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between the general education practices taught, which is likely a function of SES, which disproportionately affects minorities over Whites.
-This subtle distinction has the immediate consequence of arguing that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
+In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between general education practices used for each group, likely a function of SES, which disproportionately affects minorities.
+This subtle distinction lends credence to the arguement that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
 
 Little more can be said about the questionnaire beyond the fact that the mode for each item is the most efficient predictor of individual student responses.
-As such, we can only conclude that the information being measured by these items are independent of the exams' item functioning and student performance.
+As such, we can only conclude that the information being measured by these items is independent of the exams' item functioning and student performance.
 
@@
 
+\section{Appendix A: Explatory Factor Analysis}
+
+<<fig=T, echo=FALSE, fig.height=6, fig.width=5>>=
+fa.diagram(feng,main='English Oblimin')
+fa.diagram(fmat,main='Mathematics Oblimin')
+fa.diagram(fsci1,main='Biology Oblimin')
+@
+
 \end{document}

### 2014-03-24 paper-knitr.tex

--- 2014-03-21 paper.tex
@@
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
-\subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-11}
 
-\includegraphics{/home/landon/Documents/workspace/DataCompete/upstat-competition/figure/English.jpeg}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-12}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fmat,} \hlkwc{main} \hlstd{=} \hlstr{"Mathematics Oblimin"}\hlstd{)}
-\end{alltt}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-13}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
 \end{knitrout}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fsci1,} \hlkwc{main} \hlstd{=} \hlstr{"Biology Oblimin"}\hlstd{)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
-\end{knitrout}
+\subsection{MG-CFA}
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run three separate tests for measurement invariance:
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{e1} \hlkwb{=} \hlopt{~}\hlstd{eitem8} \hlopt{+} \hlstd{eitem10} \hlopt{+} \hlstd{eitem11} \hlopt{+} \hlstd{eitem12} \hlopt{+} \hlstd{eitem13} \hlopt{+} \hlstd{eitem16} \hlopt{+} \hlstd{eitem17} \hlopt{+} \hlstd{eitem20} \hlopt{+}
-    \hlstd{eitem21} \hlopt{+} \hlstd{eitem22} \hlopt{+} \hlstd{eitem23} \hlopt{+} \hlstd{eitem24} \hlopt{+} \hlstd{eitem25} \hlopt{+} \hlstd{eitem26} \hlopt{+} \hlstd{eitem28} \hlopt{+} \hlstd{eitem29} \hlopt{+}
-    \hlstd{eitem30} \hlopt{+} \hlstd{eitem31} \hlopt{+} \hlstd{eitem32} \hlopt{+} \hlstd{eitem33} \hlopt{+} \hlstd{eitem34} \hlopt{+} \hlstd{eitem36} \hlopt{+} \hlstd{eitem37}
-\hlstd{e2} \hlkwb{=} \hlopt{~}\hlstd{eitem27} \hlopt{+} \hlstd{eitem35} \hlopt{+} \hlstd{eitem9} \hlopt{+} \hlstd{eitem18}
-\hlstd{e3} \hlkwb{=} \hlopt{~}\hlstd{eitem38} \hlopt{+} \hlstd{eitem39} \hlopt{+} \hlstd{eitem40}
-\hlstd{erawsc} \hlopt{~} \hlstd{e1} \hlopt{+} \hlstd{e2} \hlopt{+} \hlstd{e3}
-\end{alltt}
 \begin{verbatim}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 3.981e+03 9.470e-01 9.510e-01 2.400e-02 2.154e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 4.150e+03 9.470e-01 9.490e-01 2.400e-02 2.154e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 4.337e+03 9.470e-01 9.470e-01 2.400e-02 2.154e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.eng}\hlopt{$}\hlstd{fit.residuals,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 6.008e+03 9.180e-01 9.120e-01 3.000e-02 2.169e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{m1} \hlkwb{=} \hlopt{~}\hlstd{mitem4} \hlopt{+} \hlstd{mitem5} \hlopt{+} \hlstd{mitem8} \hlopt{+} \hlstd{mitem11} \hlopt{+} \hlstd{mitem12} \hlopt{+} \hlstd{mitem14} \hlopt{+} \hlstd{mitem25} \hlopt{+} \hlstd{mitem32} \hlopt{+}
-    \hlstd{mitem35} \hlopt{+} \hlstd{mitem37} \hlopt{+} \hlstd{mitem40} \hlopt{+} \hlstd{mitem42}
-\hlstd{m2} \hlkwb{=} \hlopt{~}\hlstd{mitem6} \hlopt{+} \hlstd{mitem10} \hlopt{+} \hlstd{mitem17} \hlopt{+} \hlstd{mitem24} \hlopt{+} \hlstd{mitem39}
-\hlstd{m3} \hlkwb{=} \hlopt{~}\hlstd{mitem22} \hlopt{+} \hlstd{mitem28} \hlopt{+} \hlstd{mitem34} \hlopt{+} \hlstd{mitem38} \hlopt{+} \hlstd{mitem41}
-\hlstd{m4} \hlkwb{=} \hlopt{~}\hlstd{mitem18} \hlopt{+} \hlstd{mitem20} \hlopt{+} \hlstd{mitem21} \hlopt{+} \hlstd{mitem31} \hlopt{+} \hlstd{mitem42}
-\hlstd{mrawsc} \hlopt{~} \hlstd{m1} \hlopt{+} \hlstd{m2} \hlopt{+} \hlstd{m3} \hlopt{+} \hlstd{m4}
-\end{alltt}
 \begin{verbatim}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.configural,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 2.704e+03 9.640e-01 9.690e-01 2.800e-02 3.160e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.loadings,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 2.965e+03 9.630e-01 9.650e-01 2.900e-02 3.161e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 3.132e+03 9.630e-01 9.630e-01 2.800e-02 3.162e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.residuals,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 4.131e+03 9.490e-01 9.460e-01 3.300e-02 3.170e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{s1} \hlkwb{=} \hlopt{~}\hlstd{sitem1} \hlopt{+} \hlstd{sitem2} \hlopt{+} \hlstd{sitem13} \hlopt{+} \hlstd{sitem14} \hlopt{+} \hlstd{sitem15} \hlopt{+} \hlstd{sitem22} \hlopt{+} \hlstd{sitem27} \hlopt{+} \hlstd{sitem30} \hlopt{+}
-    \hlstd{sitem33} \hlopt{+} \hlstd{sitem34} \hlopt{+} \hlstd{sitem37} \hlopt{+} \hlstd{sitem40} \hlopt{+} \hlstd{sitem41}
-\hlstd{s2} \hlkwb{=} \hlopt{~}\hlstd{sitem3} \hlopt{+} \hlstd{sitem5} \hlopt{+} \hlstd{sitem18} \hlopt{+} \hlstd{sitem19} \hlopt{+} \hlstd{sitem28} \hlopt{+} \hlstd{sitem31} \hlopt{+} \hlstd{sitem35} \hlopt{+} \hlstd{sitem43} \hlopt{+}
-    \hlstd{sitem45}
-\hlstd{s3} \hlkwb{=} \hlopt{~}\hlstd{sitem8} \hlopt{+} \hlstd{sitem10} \hlopt{+} \hlstd{sitem11} \hlopt{+} \hlstd{sitem12} \hlopt{+} \hlstd{sitem23} \hlopt{+} \hlstd{sitem24} \hlopt{+} \hlstd{sitem25} \hlopt{+} \hlstd{sitem23}
-\hlstd{s4} \hlkwb{=} \hlopt{~}\hlstd{sitem4} \hlopt{+} \hlstd{sitem23} \hlopt{+} \hlstd{sitem29} \hlopt{+} \hlstd{sitem32} \hlopt{+} \hlstd{sitem44}
-\hlstd{srawsc} \hlopt{~} \hlstd{s1} \hlopt{+} \hlstd{s2} \hlopt{+} \hlstd{s3} \hlopt{+} \hlstd{s4}
-\end{alltt}
 \begin{verbatim}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{require}\hlstd{(mvpart)}
-\end{alltt}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-17}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: mvpart}}\begin{alltt}
-\hlkwd{mvpart}\hlstd{(}\hlkwd{cbind}\hlstd{(mrawsc, erawsc, srawsc)} \hlopt{~} \hlstd{.,} \hlkwc{data} \hlstd{= d4)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'd4' not found}}\end{kframe}
 \end{knitrout}

### 2014-03-24 paper.tex

@@
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
-\subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-11}
 
-\includegraphics{/home/landon/Documents/workspace/DataCompete/upstat-competition/figure/English.jpeg}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-12}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fmat,} \hlkwc{main} \hlstd{=} \hlstr{"Mathematics Oblimin"}\hlstd{)}
-\end{alltt}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-13}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
 \end{knitrout}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fa.diagram}\hlstd{(fsci1,} \hlkwc{main} \hlstd{=} \hlstr{"Biology Oblimin"}\hlstd{)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fa.diagram"{}}}\end{kframe}
-\end{knitrout}
+\subsection{MG-CFA}
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239)
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{fitMeasures}\hlstd{(MI.maths}\hlopt{$}\hlstd{fit.intercepts,} \hlkwd{c}\hlstd{(}\hlstr{"chisq"}\hlstd{,} \hlstr{"tli"}\hlstd{,} \hlstr{"cfi"}\hlstd{,} \hlstr{"rmsea"}\hlstd{,} \hlstr{"aic"}\hlstd{))}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}fitMeasures"{}}}\end{kframe}
+\begin{verbatim}
+##     chisq       tli       cfi     rmsea       aic 
+## 3.132e+03 9.630e-01 9.630e-01 2.900e-02 3.162e+05
+\end{verbatim}
+\end{kframe}
 \end{knitrout}

### 2014-03-24 paper-knitr.Rnw
Changes to multivariate regression code on multiple imputed questionairre data

@@
 library(mvpart)
+library(semTools)
 
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI = 0.000.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
+Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ =.000 and $\Delta$$\chi^2$=187.875.
 
@@
 \subsubsection{Strict Invariance}
-Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
+Strict invariance does not hold, with a significant $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta\mathrm{TLI}$ = .03 indicating lack of fit.
 
@@
 @
+
 The summary statistics presented below show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
+
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does not substantively change model fit above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000, the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=167.652.
+Strong Invariance does not substantively change model fit above and beyond weak invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000, the $\Delta\mathrm{TLI}$ = .000 and $\Delta$$\chi^2$=167.652.
 
@@
 \subsubsection{Strict Invariance}
-Strong Invariance imposes a significant constraint upon model fit: $\Delta$ CFI = .017, $\Delta$ RMSEA = .005, the $\Delta$ TLI = .014 and $\Delta$$\chi^2$=998.348.
+Strong Invariance imposes a significant constraint upon model fit: $\Delta\mathrm{CFI}$ = .017, $\Delta\mathrm{RMSEA}$ = .005, the $\Delta\mathrm{TLI}$ = .014 and $\Delta$$\chi^2$=998.348.
 
@@
 \subsubsection{Metric (Weak) Invariance}
-While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
+While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875. 
+Strong Invariance fails to substantively change model fit above and beyond metric invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ = .000 and $\Delta$$\chi^2$=187.875.
 <<echo=FALSE>>=
@@
 \subsubsection{Strict Invariance}
-Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
+Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006 , and the $\Delta\mathrm{TLI}$ = .03 and $\Delta$$\chi^2$=1670.297.
 <<echo=FALSE>>=

### 2014-03-24 paper-knitr.tex
Changes to multivariate regression code on multiple imputed questionairre data

@@
 
-The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
+The Massachusetts Comprehensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
 We examine a subset sample of 10\textsuperscript{th} grade students' results from the 2009 Spring MCAS test.
 The data also contains basic demographic information for each student comprised of race/ethnicity, gender, and an academic engagement questionnaire.
-Each test consists of multiple choice items and open-ended questions that are graded on a holistic scale, implemented as a 5 point integer scale.
+Each test consists of multiple choice items and open ended type questions that are graded on a holistic scale, implemented as a 5 point integer scale.
 
@@
 \item{Weak invariance:}
-Tests whether in addition to the same factor structure, the same items load equivalently onto the same same structure.
-\item{Strong invariance:}
+Tests whether in addition to the same factor structure, the same items load equivalently onto the same structure.
+\item{Strong invariance}
 Tests that the intercepts, item loadings, and factor structure are equivalent across groups.
@@
 These four factors are argued to be necessary conditions for a fair and equitable comparison\cite{Meredith}, and are established by way of multi-group confirmatory factor analysis (MG-CFA).
-In this paper, we examine these factors using a group of R packages cited at the end of the paper \cite{psych,lavaan,semTools}.
-Exploratory bootstrapped oblimin factor analyses were conducted upon each section to establish the number of domains and high loading items within each of the three tests.
+In this paper, we examine these factors using a series of R packages \cite{psych,lavaan,semTools}.
+Exploratory bootstrapped oblimin factor analyses were conducted upon each test to establish the number of domains and high loading items within each of the three tests.
 As a consequence of the extreme ethnicity imbalance within three of the four science exams, only Biology was tested using MG-CFA.
 
-Furthermore, we investigated the information found within the questionnaire, relating them to the students' scores on the three sections using both multivariate parametric regressions, and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
+Furthermore, we investigated the information found within the questionnaire by relating them to the students' scores on the three sections.
+We used both multivariate parametric regressions and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions, mainly working under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
 However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
-Even so, for the MG-CFA, we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation, in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
-However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques, producing potentially biased results under the assumption that items are missing completely at random.
-As such, we processed the original responses using the mice package\cite{mice}, using 25 imputed data frames with 30 iterations in each dataframe.
-Regression trees are relatively robust to missing data, because all information is not processed simultaneously, but sequentially in order of variable importance, and as such we did not take any additional steps when conducting that analysis.
+Even so, for the MG-CFA we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
+As such, we processed the original responses using the mice package\cite{mice} using 25 imputed data frames with 30 iterations in each dataframe.
+Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.
+As such we did not take any additional steps when conducting that analysis.
 \section{Results}
 \subsection{Exploratory factor analyses}
-Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam, using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains; factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
+Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
+Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
 The psych package was used to calculate the factor structure.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-11}
-
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-12}
-
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-13}
-
-\end{knitrout}
+Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix A.
 
 \subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run three separate tests for measurement invariance:
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239).
+Diagrams representing the EFA results can be found in Appendix A.
 
-  These were subsequently specified using a FIML procedure with bootstrapped sampling within the four primary ethnicities: A,B,H,W.
-Other ethnicities were excluded because of their relative rarity within the sample.
-As a consequence of the $\chi$\textsuperscript{2} fit statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0 \end{equation} scales as a function of both sample size and the minimum function test statistic $F_0$, traditional measurement invariance, which uses changes in significance to reject a constraint, fails here, as the sample size guarantees significant results.
-In response to this fact, we employ three different fitness criterion: the $\Delta$ Comparative Fit Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI) \cite{Chen}.
-For all three, decreases in model fit of greater than .01 are considered significant reductions in model fit.
-In addition, we include An (Akaike) Information Criterion (AIC), to demonstrate whether model fit increases or decreases between sequential steps in comparison to the theoretical true model.
+ The latent variable models for each test were specified using a FIML procedure with
+bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+Other ethnicities were excluded because of their relative rarity within
+the sample. As a consequence of the $\chi$\textsuperscript{2} fit
+statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0
+\end{equation} scales as a function of both sample size and the minimum
+function test statistic $F_0$, traditional measurement invariance, which
+uses changes in significance to reject a constraint, fails here, as the
+sample size guarantees significant results. In response to this fact, we
+employ three different fitness criterion: the $\Delta$ Comparative Fit
+Index (CFI) \cite{Bentler}, the $\Delta$ root mean square error of
+approximation (RMSEA), and the $\Delta$ Tucker-Lewis Index (TLI)
+\cite{Chen}. For all three, decreases in model fit of greater than .01
+are considered significant reductions in model fit. In addition, we
+include An (Akaike) Information Criterion (AIC), to demonstrate whether
+model fit increases or decreases between sequential steps in comparison
+to the theoretical true model.
+\subsection{English}
+The English section was constructed using the system of latent and test items presented below as predictors of the unstandardized English score.
 
-\subsection{English}
-The English section constructed using the following system of latent and test items, as predictors of the unstandardized English score, as given below.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## erawsc ~ e1 + e2 + e3
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\vspace{0.25cm}
+
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 8, 10, 11, 12, 13, 16, 17, 20,  21, 22, 23, 24, 25, 26, 28, 29,  30,
+31, 32, 33, 34, 36, 37\tabularnewline
+\hline 
+2 & 27, 35, 9, 18\tabularnewline
+\hline 
+3 & 38, 39, 40\tabularnewline
+\hline
+\end{tabular}
+
+\bigskip
+
+All latent variable indicators are questions from the English section of the test (e.g. `8' represents `eitem8').
+
+\end{minipage}
 
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI =.000 and $\Delta$$\chi^2$=187.875.
+Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ =.000 and $\Delta$$\chi^2$=187.875.
+
 \begin{knitrout}
@@
 \subsubsection{Strict Invariance}
-Strict invariance does not hold, with a significant $\Delta$ CFI = .034, $\Delta$ RMSEA = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta$ TLI = .03 indicating lack of fit.
+Strict invariance does not hold, with a significant $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta\mathrm{TLI}$ = .03 indicating lack of fit.
+
 \begin{knitrout}
@@
 \subsection{Mathematics}
-While the EFA presented a good model fit for a five factor model, when the SEM model imposed upon the data, it failed to converge due to underspecification.
-As such, factor 5 was removed, because it consisted of only four low loading items.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## mrawsc ~ m1 + m2 + m3 + m4
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+While the EFA presented a good model fit for a five factor model, when the SEM model was imposed upon the data it failed to converge due to underspecification.
+As such, factor 5 was removed because it consisted of only four low loading items.
+The latent variable model for Mathematics is presented in the table below.
+
+\vspace{0.25cm}
+
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 4, 5, 8, 11, 12, 14, 25, 32, 35, 37, 40, 42\tabularnewline
+\hline 
+2 & 6, 10, 17, 24, 39\tabularnewline
+\hline 
+3 & 22, 28, 34, 38, 41\tabularnewline
+\hline 
+4 & 18, 20, 21, 31, 42\tabularnewline
+\hline 
+\end{tabular}
+
+\bigskip
+
+All latent variable indicators are questions from the Mathematics section of the test (e.g. `8' represents `mitem8').
+
+\end{minipage}
 
@@
 
-As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
+The summary statistics presented below show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
+
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .004, $\Delta$ RMSEA = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta$ TLI = .002.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
 
@@
 ##     chisq       tli       cfi     rmsea       aic 
-## 3.132e+03 9.630e-01 9.630e-01 2.800e-02 3.162e+05
+## 3.132e+03 9.630e-01 9.630e-01 2.900e-02 3.162e+05
 \end{verbatim}
@@
 In addition, each subsequent test produced similar results, indicating that across all four primary ethnicities, the test operated with equivalent measurement.
-Biology as presented an interpretable factor structure, which was represented with the following mathematical structure.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## srawsc ~ s1 + s2 + s3 + s4
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+Biology presented an interpretable factor structure, which was represented with the following mathematical structure.
+
+\vspace{0.25cm}
+
+\begin{minipage}{\linewidth}
+\begin{tabular}{|c|l|}
+\multicolumn{1}{c}{Latent Variable} & \multicolumn{1}{l}{Latent Variable Indicators}\tabularnewline
+\hline 
+1 & 1, 2, 13, 14, 15, 22, 27, 30, 33, 34, 37, 40, 41\tabularnewline
+\hline 
+2 & 3, 5, 18, 19, 28, 31, 35, 43, 45\tabularnewline
+\hline 
+3 & 8, 10, 11, 12, 23, 24, 25\tabularnewline
+\hline 
+4 & 4, 23, 29, 32, 44\tabularnewline
+\hline 
+\end{tabular}
+
+\bigskip
+
+All latent variable indicators are questions from the Biology section of the test (e.g. `8' represents `sitem8').
+
+\end{minipage}
 
 \subsubsection{Configural Invariance}
+The analyses' summary statistics show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
+
 \begin{knitrout}
@@
 
-As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items. 
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta$ TLI = .000.
+While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
 
@@
 \subsubsection{Strong Invariance}
-Strong Invariance does fails to substantively change model fit above and beyond metric invariance: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , and the $\Delta$ TLI = .000 and $\Delta$$\chi^2$=187.875.
+Strong Invariance fails to substantively change model fit above and beyond metric invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ = .000 and $\Delta$$\chi^2$=187.875.
 \begin{knitrout}
@@
 \subsubsection{Strict Invariance}
-Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta$ CFI = .034, $\Delta$ RMSEA = .006 , and the $\Delta$ TLI = .03 and $\Delta$$\chi^2$=1670.297. 
+Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006 , and the $\Delta\mathrm{TLI}$ = .03 and $\Delta$$\chi^2$=1670.297.
 \begin{knitrout}
@@
 \subsubsection{Multivariate Regression Tree}
-
-Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath}, conceptually similar to running a multinomial regression upon the different clusters analysis groups of the response variables.
+Without preexisting beliefs about the nature of the relationship between the response surface and the questionnaire and demographics, we chose to begin the analyses with an undirected multivariate tree regression \cite{De'ath} which is conceptually similar to running a multinomial regression upon the different clusters of the response variables.
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-17}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-13}
 
@@
 These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
-The failure for the function to further discretize the low achievers indicates that the individual differences wihin accounted for no additional information.
-For high aspiring students, ethnicity was found to be the second most important in discriminating test scores, with Blacks, Hispanics and Native Americans performing similarly worse on the Mathematics and Biology sections than the other ethnicities.
+The failure for the function to further discretize the low achievers indicates that the individual differences within this group accounted for no additional information.
+For high aspiring students, ethnicity was found to be the second most important variable in discriminating test scores: Blacks, Hispanics and Native Americans performed similarly worse on the Mathematics and Biology sections compared to the other ethnicities.
 The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.
 Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15
+
 \subsubsection{Multivariate Regression}
-A parametric regression model was also conducted exploring only main effects between the predictors and response.
+A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
 
 \section{Discussion}
-As indicated in the results section, the test fails to perform equivalently across the four groups, which is the primary point of order in test design and advocacy.
-Moreover, strict invariance has a specific import upon the function of psychometric and education measurements, testing whether the error variances --the portions of item variation not attributable to the common factors.
-Conceptually, strict invariance validates the assumption that error variances are different across groups; if this fails, there are either different variables operating on the measures across groups or the same set of variables operates differently across groups \cite{deshon}.
-This fundamentally demonstrates the importance of ensuring that all four mainstays of measurement invariance be substantiated, especially when one considers the long standing view that residuals convey meaning of unmeasured latent constructs' effects, moderated by group membership in this case \cite{cronbach}.
-This lends substantive credence to the argument that there are fundamental flaws within the test, specifically with reflection towards students' socioeconomic status.
+
+The primary point of order in test design is for a test to perform equivalently across groups of testees.
+As indicated in the results section, our analysis suggests that the test fails to perform equivalently across four ethnicity groups.
+This supplies quantitative evidence to the argument that there are fundamental flaws within the test.
+However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
+In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between general education practices used for each group, likely a function of SES, which disproportionately affects minorities.
+This subtle distinction lends credence to the arguement that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
+
+Little more can be said about the questionnaire beyond the fact that the mode for each item is the most efficient predictor of individual student responses.
+As such, we can only conclude that the information being measured by these items is independent of the exams' item functioning and student performance.
 
@@
 
+\section{Appendix A: Explatory Factor Analysis}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-141}
+
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-142}
+
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-143}
+
+\end{knitrout}
+
 \end{document}

### 2014-03-24 EDA work.R

@@ -167,11 +167,16 @@ d5$race_off<-NULL
 require(mice)
 d6<-mice(d5,m=13,maxit=15,defaultMethod = c("norm","logreg","polyreg"))
 summary(aggr(d5[,c(1,7:31)]))
-summary(manova(cbind(mrawsc,erawsc,srawsc)~.,d5)) #this is on the data with missing 
-summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5)) #this is on the data with missing 
 
-summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~.)))
-summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~.)))
+d5.fixed = subset(d5, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
+
+summary(manova(cbind(mrawsc,erawsc,srawsc)~.,d5.fixed)) #this is on the data with missing 
+summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5.fixed)) #this is on the data with missing 
+
+#d6.fixed = subset(d6$imp, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
+
+summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~., d6.fixed)))
+summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~., d6.fixed)))
 
 # require(VIM)
 # a$erawsc<-a$emcpts<-a$eorpts<-a$ecpi<-a$mrawsc<-a$mmcpts<-a$morpts<-a$mcpi<-a$srawsc<-a$smcpts<-a$sorpts<-a$scpi<-a$eitem41<-a$eitem42<-a$scitry<-NULL

### 2014-03-24 EDA work.R
Changes to multivariate regression code on multiple imputed questionairre data

@@ -175,8 +175,8 @@ summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5.fixed)) #this is on the data with mi
 
 #d6.fixed = subset(d6$imp, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
 
-summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~., d6.fixed)))
-summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~., d6.fixed)))
+summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~., d6)))
+summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~., d6)))
 
 # require(VIM)
 # a$erawsc<-a$emcpts<-a$eorpts<-a$ecpi<-a$mrawsc<-a$mmcpts<-a$morpts<-a$mcpi<-a$srawsc<-a$smcpts<-a$sorpts<-a$scpi<-a$eitem41<-a$eitem42<-a$scitry<-NULL

### 2014-03-24 paper-knitr.Rnw
updates

@@
 We performed a series of statistical analyses on a subset of 10\textsuperscript{th} grade student test results in English, Mathematics, and Biology tests.
-An important property of any testing procedure is that of measurement invariance: the test should measure the same underlying constructs, in the same manner, regardless of what groups (ethnicicy, gender, etc.) each testee belongs to.
+An important property of any testing procedure is that of measurement invariance: the test should measure the same underlying constructs, in the same manner, regardless of what groups (ethnicity, gender, etc.) each testee belongs to.
 We employed Structural Equation Modeling (SEM) to test for measurement invariance across four ethnic groups (White, Black, Hispanic, and Asian) in the MCAS results.
@@
 The former, comprised mainly of teachers and educational policy analysts, claim that the test is unfairly biased against students as a function of demographics, while citing the significant impediment that poor performance upon the tests has upon both immediate and long-term outcomes.
-\cite{Gaudet} Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
+\cite{Gaudet} Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealistic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
 
 \subsection{Theoretical structure of the test}
-Our paper begins with an analysis of the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 
@@
 
-Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relaionship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
+Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relationship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
 
 \subsection{Procedure}
-A critical assumption within standardised testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
+A critical assumption within standardized testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
 This property is a series of mathematical constraints applied to latent variables that are assumed to be non-significantly different between groups.
-Latent variables are unobservable constructs that underly the observed measures; one typically estimates them by computing standard principal axis factor analysis.
+Latent variables are unobservable constructs that underlie the observed measures; one typically estimates them by computing standard principal axis factor analysis.
 The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample.
@@
 \item{Strict invariance:}
-Tests the assumption that in addition to the preceeding steps, the residual variances are equivalent across groups.
+Tests the assumption that in addition to the preceding steps, the residual variances are equivalent across groups.
 \end{enumerate}
@@
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
-However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
-Even so, for the MG-CFA we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+However, within the structure of the exam, missing data was unsurprisingly a rare occurrence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
+Even so, for the MG-CFA we attempted to account for this by utilizing the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
 However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
-As such, we processed the original responses using the mice package\cite{mice} using 25 imputed data frames with 30 iterations in each dataframe.
+As such, we processed the original responses using the mice package\cite{mice} using 13 imputed data frames with 15 iterations in each dataframe.
 Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.
@@
 \subsection{Exploratory factor analyses}
-Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
+Given the preceding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
 Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
 
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
 
@@
 \subsubsection{Metric (Weak) Invariance}
-While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
+While Weak Invariance does impact the model fit compared to step 1, the change is negligible: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
 
@@
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
-We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate nonresponse as a partitioning criterion.
+We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate non-response as a partitioning criterion.
 
@@
 \subsubsection{Multivariate Regression}
-A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
+A parametric regression model was also conducted exploring only main effects between the predictors and response.
+Secondary analyses run upon the data, using the VIM package, clearly indicated that the missing data was uniformly equal for all but one question, which substantively improves the likelihood that these questions are truly missing completely at random (MCAR).
+In addition, the bulk of the missing information was a result of all questions being skipped together.
+Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical, thus never distributed normally.
+All tests results were based on both the imputed and missing data, however no differences were found amongst the missing results.
+Results were singly focused upon the science sections, with more involved educational settings supporting higher outcomes.
+Males typically outperformed women, and Hispanic/Latino and Black students performed significantly worse than other ethnicity groups, who performed non-significantly differentially from themselves.
+Otherwise, student achievement outcomes were the only conclusively meaningful predictors that presented, with students who expected to achieve higher (e.g., achieve a baccalaureate degree) having higher outcomes.
+All code necessary to replicate these effects are submitted within the R script file submitted with these results.
 
@@
 This supplies quantitative evidence to the argument that there are fundamental flaws within the test.
-However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surpising given the previous point.
+However, the multivariate regression and tree results indicate that there are no functional differences between the test scores themselves across ethnicity, which is surprising given the previous point.
 In our opinion, a plausible explanation for these results are that there are ethnic differences not between the overall scores but between general education practices used for each group, likely a function of SES, which disproportionately affects minorities.
-This subtle distinction lends credence to the arguement that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
+This subtle distinction lends credence to the argument that there is a fundamental flaw within the standards of difficulty construed as appropriate for this test, and as a consequence, the test performs poorly as an indicator of individual student achievement.
 
@@
 
-\section{Appendix A: Explatory Factor Analysis}
-
-<<fig=T, echo=FALSE, fig.height=6, fig.width=5>>=
-fa.diagram(feng,main='English Oblimin')
-fa.diagram(fmat,main='Mathematics Oblimin')
-fa.diagram(fsci1,main='Biology Oblimin')
-@
-
 \end{document}

### 2014-03-24 paper-knitr.tex
updates

@@
 However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
-As such, we processed the original responses using the mice package\cite{mice} using 25 imputed data frames with 30 iterations in each dataframe.
+As such, we processed the original responses using the mice package\cite{mice} using 13 imputed data frames with 15 iterations in each dataframe.
 Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.
@@
 \subsubsection{Multivariate Regression}
-A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
+A parametric regression model was also conducted exploring only main effects between the predictors and response.
+Secondary analyses run upon the data, using the VIM package, clearly indicated that the missing data was uniformly equal for all but one question, which substantively improves the likelihood that these questions are truly missing completely at random (MCAR).
+In addition, the bulk of the missing information was a result of all questions being skipped together.
+Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical, thus never distributed normally.
+All tests results were based on both the imputed and missing data, however no differences were found amongst the missing results.
+Results were singly focused upon the science sections, with more involved educational settings supporting higher outcomes.
+Males typically outperformed women, and Hispanic/Latino and Black students performed significantly worse than other ethnicity groups, who performed non-significantly differentally from themselves.
+Otherwise, student achievement outcomes were the only conclusively meaningful outcomes that presented, with students who expected to achieve higher having higher outcomes.
 
@@
 
-\section{Appendix A: Explatory Factor Analysis}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-141}
-
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-142}
-
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-143}
-
-\end{knitrout}
-
 \end{document}

### 2014-03-24 EDA work.R

@@ -154,6 +154,8 @@ fitMeasures(MI.scien$fit.means, c('chisq','tli', 'cfi','rmsea','aic'))
 d4<-d[,c(145:172,45,91,141,95,2)]
 require(mvpart)
 fit<-mvpart(cbind(mrawsc,erawsc,srawsc)~.,d4)
+summary(fit)
+
 
 require(foreign)
 a<-read.spss('/home/landon/Documents/workspace/DataCompete/Student Data Competition/missingDataCoded.sav',use.value.labels=T,to.data.frame=T,use.missings=T)
@@ -170,13 +172,13 @@ summary(aggr(d5[,c(1,7:31)]))
 
 d5.fixed = subset(d5, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
 
-summary(manova(cbind(mrawsc,erawsc,srawsc)~.,d5.fixed)) #this is on the data with missing 
-summary(lm(cbind(mrawsc,erawsc,srawsc)~.,d5.fixed)) #this is on the data with missing 
+summary(manova(cbind(mrawsc,erawsc,srawsc)~.+eth*gender,d5.fixed)) #this is on the data with missing 
+summary(lm(cbind(mrawsc,erawsc,srawsc)~.+eth*gender,d5.fixed)) #this is on the data with missing 
 
 #d6.fixed = subset(d6$imp, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
 
-summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~., d6)))
-summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~., d6)))
+summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~.+eth*gender, d6)))
+summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~.+eth*gender, d6)))
 
 # require(VIM)
 # a$erawsc<-a$emcpts<-a$eorpts<-a$ecpi<-a$mrawsc<-a$mmcpts<-a$morpts<-a$mcpi<-a$srawsc<-a$smcpts<-a$sorpts<-a$scpi<-a$eitem41<-a$eitem42<-a$scitry<-NULL

### 2014-03-24 paper-knitr.Rnw
updates

--- 2014-03-24 paper-knitr.Rnw
@@
 \subsubsection{Multivariate Regression}
-A parametric regression model was also conducted exploring only main effects between the predictors and response. Multivariate and univariate responses were found within the tests
+A parametric regression model was also conducted exploring only main effects between the predictors and response.
+Secondary analyses run upon the data, using the VIM package, clearly indicated that the missing data was uniformly equal for all but one question, which substantively improves the likelihood that these questions are truly missing completely at random (MCAR).
+In addition, the bulk of the missing information was a result of all questions being skipped together.
+Unfortunately, more algorithmically driven tests \cite{little,KB,}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical, thus never distributed normally.
+All tests results were based on both the imputed and missing data, however no differences were found amongst the missing results.
+Results were singly focused upon the science sections, with more involved educational settings supporting higher outcomes.
+Males typically outperformed women, and Hispanic/Latino and Black students performed significantly worse than other ethnicity groups, who performed non-significantly differentally from themselves.
+Otherwise, student achievement outcomes were the only conclusively meaningful outcomes that presented, with students who expected to achieve higher having higher outcomes.
 

### 2014-03-24 paper-knitr.Rnw
updates3

--- 2014-03-24 paper-knitr.Rnw
@@
 As such we did not take any additional steps when conducting that analysis.
+=======
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
+Even so, for the MG-CFA we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
+As such, we processed the original responses using the mice package\cite{mice} using 13 imputed data frames with 15 iterations in each dataframe.
+Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.
+As such we did not take any additional steps when conducting that analysis.
+>>>>>>> 782b5472c392dad1d398b23760e87d19f21ebe4a
 \section{Results}
@@
 \subsubsection{Multivariate Regression}
+
 A parametric regression model was also conducted exploring only main effects between the predictors and response.

### 2014-03-25 paper-knitr.Rnw
Changes per Dr. Bilgic's recommendations

--- 2014-03-24 paper-knitr.Rnw
@@
 
-\subsection{Theoretical structure of the test}
 Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 
-\section{Methods}
+\section{Methodology}
 \subsection{Subjects}
@@
 
-\subsection{Procedure}
+\subsection{Test Validity}
 A critical assumption within standardized testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
@@
 
-Furthermore, we investigated the information found within the questionnaire by relating them to the students' scores on the three sections.
+\subsection{Questionairre Results}
+
+We investigated the information found within the student questionnaire by relating them to the students' scores on the three sections.
 We used both multivariate parametric regressions and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
@@
 As such we did not take any additional steps when conducting that analysis.
+
 \section{Results}
@@
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix A.
+Items that loaded higher than .3 on a factor were specified as indicators.
+See Appendix A for diagrams of the factor structures.
 
@@
  The latent variable models for each test were specified using a FIML procedure with
-bootstrapped sampling within the four primary ethnicities: A,B,H,W.
+bootstrapped sampling within the four primary ethnicities: A, B, H, W.
 Other ethnicities were excluded because of their relative rarity within
-the sample. As a consequence of the $\chi$\textsuperscript{2} fit
-statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) * F_0
+the sample. The $\chi^{2}$ goodness of fit statistic was used to evaluate the fit of the model. However, this statistic, defined as \begin{equation} \chi^2_{d.f}=2(N-1) \cdot F_0
 \end{equation} scales as a function of both sample size and the minimum
 function test statistic $F_0$, traditional measurement invariance, which
-uses changes in significance to reject a constraint, fails here, as the
+uses changes in significance to reject a constraint, fails here as the
 sample size guarantees significant results. In response to this fact, we
@@
 are considered significant reductions in model fit. In addition, we
-include An (Akaike) Information Criterion (AIC), to demonstrate whether
+include An Akaike Information Criterion (AIC), to demonstrate whether
 model fit increases or decreases between sequential steps in comparison
@@
 
-All latent variable indicators are questions from the English section of the test (e.g. `8' represents `eitem8').
+All latent variable indicators are questions from the English section of the test (e.g. `8' represents Question 8, column `eitem8' in the provided dataset).
 
@@
 
-\subsubsection{Configural Invariance}
+\subsubsection*{Configural Invariance}
 <<echo=FALSE>>=
@@
 As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
-\subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
+\subsubsection*{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to Configural Invariance, however negligibly: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
 
@@
 
-\subsubsection{Strong Invariance}
+\subsubsection*{Strong Invariance}
 Strong Invariance is established using the model fit indexes above and beyond weak invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ =.000 and $\Delta$$\chi^2$=187.875.
@@
 
-\subsubsection{Strict Invariance}
+\subsubsection*{Strict Invariance}
 Strict invariance does not hold, with a significant $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006, $\Delta$$\chi^2$=1670.297, and the $\Delta\mathrm{TLI}$ = .03 indicating lack of fit.
@@
 
-All latent variable indicators are questions from the Mathematics section of the test (e.g. `8' represents `mitem8').
+All latent variable indicators are questions from the Mathematics section of the test (e.g. `8' represents Question 8, column `mitem8' in the provided dataset).
 
@@
 
-\subsubsection{Metric (Weak) Invariance}
+\subsubsection*{Metric (Weak) Invariance}
 Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
@@
 
-\subsubsection{Strong Invariance}
+\subsubsection*{Strong Invariance}
 Strong Invariance does not substantively change model fit above and beyond weak invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000, the $\Delta\mathrm{TLI}$ = .000 and $\Delta$$\chi^2$=167.652.
@@
 
-\subsubsection{Strict Invariance}
+\subsubsection*{Strict Invariance}
 Strong Invariance imposes a significant constraint upon model fit: $\Delta\mathrm{CFI}$ = .017, $\Delta\mathrm{RMSEA}$ = .005, the $\Delta\mathrm{TLI}$ = .014 and $\Delta$$\chi^2$=998.348.
@@
 
-All latent variable indicators are questions from the Biology section of the test (e.g. `8' represents `sitem8').
+All latent variable indicators are questions from the Biology section of the test (e.g. `8' represents Question 8, column `sitem8' in the provided dataset).
 
@@
 
-\subsubsection{Configural Invariance}
+\subsubsection*{Configural Invariance}
 The analyses' summary statistics show that configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
@@
 
-\subsubsection{Metric (Weak) Invariance}
+\subsubsection*{Metric (Weak) Invariance}
 While Weak Invariance does impact the model fit compared to step 1, the change is negligible: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
@@
 
-\subsubsection{Strong Invariance}
+\subsubsection*{Strong Invariance}
 Strong Invariance fails to substantively change model fit above and beyond metric invariance: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , and the $\Delta\mathrm{TLI}$ = .000 and $\Delta$$\chi^2$=187.875.
@@
 
-\subsubsection{Strict Invariance}
+\subsubsection*{Strict Invariance}
 Strict Invariance does substantively change model fit above and beyond strong invariance: $\Delta\mathrm{CFI}$ = .034, $\Delta\mathrm{RMSEA}$ = .006 , and the $\Delta\mathrm{TLI}$ = .03 and $\Delta$$\chi^2$=1670.297.
@@
 
+\begin{figure}[h]
 <<fig=T, echo=FALSE, fig.height=4>>=
@@
 @
+\caption{Multivariate regression tree of questionnaire results.
+\emph{`mrawsc'}, \emph{`erawsc'}, and \emph{`srawc'} are the total Mathematics, English, and Biology scores for each student, respectively.
+\emph{`questX'} is questionnaire question X, and \emph{`eth'} is the reported ethnicity.
+}
+\end{figure}
 
@@
 The final node split discriminates between high and low scoring science students, with those who responded as rarely using scientific instruments performing worse.
-Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15
+Cross-validated errors were high, at CVE=.905, which corresponds to $R^2$=.15.
 
 \subsubsection{Multivariate Regression}
+
 A parametric regression model was also conducted exploring only main effects between the predictors and response.
@@
 In addition, the bulk of the missing information was a result of all questions being skipped together.
-Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical, thus never distributed normally.
+Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical.
 All tests results were based on both the imputed and missing data, however no differences were found amongst the missing results.
@@
 All code necessary to replicate these effects are submitted within the R script file submitted with these results.
+
+The regression tree can capture interactions between variables and nonlinear effects, while the multivariate regression performed only measures linear effects.
+The fact that the $R^2$ values for each analysis are similar suggests that there may not be significant interactions or nonlinear effects in the relationship between questionnaire results and test scores.
 
@@
 
+\section{Appendix A: Exploratory Factor Analysis}
+
+<<fig=T, echo=FALSE, fig.height=6, fig.width=5>>=
+fa.diagram(feng,main='English Oblimin')
+fa.diagram(fmat,main='Mathematics Oblimin')
+fa.diagram(fsci1,main='Biology Oblimin')
+@
+
 \end{document}

### 2014-03-25 paper-knitr.tex
updates
Changes per Dr. Bilgic's recommendations

@@
 The former, comprised mainly of teachers and educational policy analysts, claim that the test is unfairly biased against students as a function of demographics, while citing the significant impediment that poor performance upon the tests has upon both immediate and long-term outcomes.
-\cite{Gaudet} Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealitic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
+\cite{Gaudet} Theoretical perspectives on test design are the primary domain of psychometrics, who hold the unfortunate distinction of both supporting standards-based assessment, and subsequently watching their recommendations be ignored when put into practice, either through misinterpretation of statistical results, or the implementation of unrealistic expectations that ignore the original purpose: measurement, or the designation of numerical scales to classify underlying unobservable latent constructs.
 
 \subsection{Theoretical structure of the test}
-Our paper begins with an analysis of the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 
-\section{Methods}
+\section{Methodology}
 \subsection{Subjects}
@@
 
-Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relaionship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
+Unfortunately for our purposes, the primary point of contention regarding exam disparity, student socio-economic status (SES), was not available, which limits our ability to test and explore the purported causal relationship with student scores \cite{Gaudet}, as mediated by ethnicity or spatial location.
 
-\subsection{Procedure}
-A critical assumption within standardised testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
+\subsection{Test Validity}
+A critical assumption within standardized testing, and in truth any measurement scale, is that the test demonstrates a psychometric property called measurement invariance/equivalence.
 This property is a series of mathematical constraints applied to latent variables that are assumed to be non-significantly different between groups.
-Latent variables are unobservable constructs that underly the observed measures; one typically estimates them by computing standard principal axis factor analysis.
+Latent variables are unobservable constructs that underlie the observed measures; one typically estimates them by computing standard principal axis factor analysis.
 The constraints serve to empirically demonstrate that across sub-populations, the test measures the same concepts, and that each item tests in a similar way the same construct in comparison to the overall sample.
@@
 
-Furthermore, we investigated the information found within the questionnaire by relating them to the students' scores on the three sections.
+\subsection{Questionairre Results}
+
+We investigated the information found within the student questionnaire by relating them to the students' scores on the three sections.
 We used both multivariate parametric regressions and a non-parametric predictive procedure using the mvpart package, which models multivariate regression trees.
 
-Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
-However, within the structure of the exam, missing data was unsurprisingly a rare occurence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
-Even so, for the MG-CFA we attempted to account for this by utilising the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
+Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}.
+However, within the structure of the exam, missing data was unsurprisingly a rare occurrence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions.
+Even so, for the MG-CFA we attempted to account for this by utilizing the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors.
 However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
@@
 As such we did not take any additional steps when conducting that analysis.
+
 \section{Results}
 \subsection{Exploratory factor analyses}
-Given the preceeding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
+Given the preceding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains.
 Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items.
@@
 
-\subsubsection{Configural Invariance}
+\subsubsection*{Configural Invariance}
 \begin{knitrout}
@@
 As the summary statistics given in the analyses show, configural invariance is supported, meaning that all four ethnicities employ the same conceptual frameworks to answer the test items.
-\subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
+\subsubsection*{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to Configural Invariance, however negligibly: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = 0.000 , $\Delta$$\chi^2$=187.875, and the $\Delta\mathrm{TLI}$ = 0.000.
 
@@
 
-\subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
+\subsubsection*{Metric (Weak) Invariance}
+Weak Invariance does impact the model fit compared to step 1, however negligibly: $\Delta\mathrm{CFI}$ = .004, $\Delta\mathrm{RMSEA}$ = .001 , $\Delta$$\chi^2$=260.076, and the $\Delta\mathrm{TLI}$ = .002.
 
@@
 
-\subsubsection{Metric (Weak) Invariance}
-While Weak Invariance does impact the model fit compared to step 1, the change is negligable: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
+\subsubsection*{Metric (Weak) Invariance}
+While Weak Invariance does impact the model fit compared to step 1, the change is negligible: $\Delta\mathrm{CFI}$ = .002, $\Delta\mathrm{RMSEA}$ = .000 , $\Delta$$\chi^2$=168.330, and the $\Delta\mathrm{TLI}$ = .000.
 
@@
 The tree also tests by default the possibility that which science test a student takes interacts with other outcomes --in addition, forcibly partitioning the data does not meaningfully increase model fit.
-We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate nonresponse as a partitioning criterion.
+We also made the decision to leave non-response as a legitimate factor level: this allows us to incorporate non-response as a partitioning criterion.
 
+\begin{figure}[h]
 \begin{knitrout}
@@
 
+\caption{Multivariate regression tree of questionnaire results.
+\emph{`mrawsc'}, \emph{`erawsc'}, and \emph{`srawc'} are the total Mathematics, English, and Biology scores for each student, respectively.
+\emph{`questX'} is questionnaire question X, and \emph{`eth'} is the reported ethnicity.
+}
+\end{figure}
+
 These results indicate substantive effects between the future student aspirations (those who intended to attain a baccalaureate degree) and overall performance, with these students performing significantly better on all three sections.
@@
 In addition, the bulk of the missing information was a result of all questions being skipped together.
-Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical, thus never distributed normally.
+Unfortunately, more algorithmically driven tests \cite{little,KB}, traditionally require that the data in question be normally distributed, and then proceed with homoscedatic assessments of the data; for our purposes, the predictors are all categorical.
 All tests results were based on both the imputed and missing data, however no differences were found amongst the missing results.
 Results were singly focused upon the science sections, with more involved educational settings supporting higher outcomes.
-Males typically outperformed women, and Hispanic/Latino and Black students performed significantly worse than other ethnicity groups, who performed non-significantly differentally from themselves.
-Otherwise, student achievement outcomes were the only conclusively meaningful outcomes that presented, with students who expected to achieve higher having higher outcomes.
+Males typically outperformed women, and Hispanic/Latino and Black students performed significantly worse than other ethnicity groups, who performed non-significantly differentially from themselves.
+Otherwise, student achievement outcomes were the only conclusively meaningful predictors that presented, with students who expected to achieve higher (e.g., achieve a baccalaureate degree) having higher outcomes.
+All code necessary to replicate these effects are submitted within the R script file submitted with these results.
+
+The regression tree can capture interactions between variables and nonlinear effects, while the multivariate regression performed only measures linear effects.
+The fact that the $R^2$ values for each analysis are similar suggests that there may not be significant interactions or nonlinear effects in the relationship between questionnaire results and test scores.
 
@@
 
+\section{Appendix A: Exploratory Factor Analysis}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-141}
+
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-142}
+
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-143}
+
+\end{knitrout}
+
 \end{document}

### 2014-03-25 paper-knitr.Rnw
Passed through a spellchecker

@@
 
-\title{An Analysis of Massachusetts' Standardized Testing through Multi-Group Structural Equation Modelling}
+\title{An Analysis of Massachusetts' Standardized Testing through Multi-Group Structural Equation Modeling}
 \date{March 24, 2014}
@@
 \section{Introduction}
-
-\subsection{Data}
 
@@
 
-\subsection{Questionairre Results}
+\subsection{Questionnaire Results}
 
@@
 However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random.
-As such, we processed the original responses using the mice package\cite{mice} using 13 imputed data frames with 15 iterations in each dataframe.
+As such, we processed the original responses using the mice package\cite{mie} using 13 imputed data frames with 15 iterations in each dataframe.
 Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance.

### 2014-03-25 EDA work.R
mice fixes, if you want to try subsetting the regression coefficients.

@@ -171,15 +171,25 @@ d6<-mice(d5,m=13,maxit=15,defaultMethod = c("norm","logreg","polyreg"))
 summary(aggr(d5[,c(1,7:31)]))
 
 d5.fixed = subset(d5, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
-
+options(contrasts = c("contr.sum", "contr.poly"))
 summary(manova(cbind(mrawsc,erawsc,srawsc)~.+eth*gender,d5.fixed)) #this is on the data with missing 
 summary(lm(cbind(mrawsc,erawsc,srawsc)~.+eth*gender,d5.fixed)) #this is on the data with missing 
 
 #d6.fixed = subset(d6$imp, select=-c(quest2a, quest2b, quest2c, quest2d, quest2e))
 
-summary(with(d6,manova(cbind(mrawsc,erawsc,srawsc)~.+eth*gender, d6)))
-summary(with(d6,lm(cbind(mrawsc,erawsc,srawsc)~.+eth*gender, d6)))
 
+
+
+w1=with(d6,lm(erawsc~quest1+quest3+quest4+quest5+quest6+quest7+quest8+quest9+quest10+quest11+quest12+quest13+quest14+quest15+quest16+quest17+quest18+quest19+quest20+quest21+gender*eth))
+w2=with(d6,lm(mrawsc~quest1+quest3+quest4+quest5+quest6+quest7+quest8+quest9+quest10+quest11+quest12+quest13+quest14+quest15+quest16+quest17+quest18+quest19+quest20+quest21+gender*eth))
+w3=with(d6,lm(srawsc~quest1+quest3+quest4+quest5+quest6+quest7+quest8+quest9+quest10+quest11+quest12+quest13+quest14+quest15+quest16+quest17+quest18+quest19+quest20+quest21+gender*eth))
+
+pool.r.squared(w1,adjusted=T)
+pool.r.squared(w2,adjusted=T)
+pool.r.squared(w3,adjusted=T)
+round(summary(pool(w1))[,c(1,3,5)],3)
+round(summary(pool(w2))[,c(1,3,5)],3)
+round(summary(pool(w3))[,c(1,3,5)],3)
 # require(VIM)
[…]

### 2014-04-09 presentation.Rnw
Presentation beamer file

\documentclass{beamer}

\title{Analysis of Massachusetts' Standardized Testing Multi-Group Structural Equation Modelling}
\author[shortname]{Landon Hurley \inst{1} \and Herb Susmann \inst{2}}
\institute[shortinst]{\inst{1} Psychology, SUNY Geneseo \and %
\inst{2} Mathematics, SUNY Geneseo}
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \tableofcontents
\end{frame}

\begin{frame}{Introduction}

\end{frame}

\section{Data}

  \begin{frame}{Data}
    \begin{itemize}

    \end{itemize}
  \end{frame}

\section{Measurement Invariance}

  \begin{frame}{Measurement Invariance}

  \end{frame}

  \subsection{Results}

    \begin{frame}{Results}

    \end{frame}

  \subsection{Discussion}

    \begin{frame}{Discussion}

    \end{frame}

\section{Regression Tree}
  \begin{frame}{Regression Tree}

  \end{frame}

  \subsection{Results}

    \begin{frame}{Results}

    \end{frame}

  \subsection{Discussion}

    \begin{frame}{Discussion}

    \end{frame}

\end{document}

### 2014-04-11 paper.Rnw
Presentation updates

@@
 \subsection{Theoretical structure of the test}
-Our paper begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigration into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
+Our paper begins with an analaysis of both the psychometric equivalance of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.
 \section{Methods}
@@
 The psych package was used to calculate the factor structure.
-Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix B.
-\subsection{MG-CFA}
-As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239); exploratory factor analyses (EFA) presented the following results:
-
-\includegraphics{/home/landon/Documents/workspace/DataCompete/upstat-competition/figure/English.jpeg}
-
-<<fig=T>>=
+<<fig=T,echo=F>>=
+fa.diagram(feng,main='English Oblimin')
 fa.diagram(fmat,main='Mathematics Oblimin')
-@
-
-<<fig=T>>=
 fa.diagram(fsci1,main='Biology Oblimin')
 @
+\subsection{MG-CFA}
+As a consequence of the unequal numbers of test takers between Biology and the rest of the science exam options, we chose to run two separate tests for measurement invariance: the first was composed of the total sample on the English and Mathematics, and the second was composed of the Biology testees ($N$=9,239)
 
@@
 \subsubsection{Metric (Weak) Invariance}
-Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA =. , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
+Weak Invariance does impact the model fit compared to step 1, however negligably: $\Delta$ CFI = .002, $\Delta$ RMSEA = .000 , $\Delta$$\chi^2$=187.875, and the $\Delta$ TLI =.
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.maths$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<>>=
-fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic'))
+<<echo=F>>=
+round(fitMeasures(MI.maths$fit.intercepts, c('chisq','tli', 'cfi','rmsea','aic')),3)
 @
@@
 
-<<>>=
+<<echo=F>>=
 fitMeasures(MI.eng$fit.loadings, c('chisq','tli', 'cfi','rmsea','aic'))
@@
 
-<<fig=T>>=
-require(mvpart)
+<<echo=F,fig=T>>=
 mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4)

### 2014-04-11 presentation.Rnw
Presentation updates

@@
 \documentclass{beamer}
+
+<<setup, include=F, cache=FALSE, echo=FALSE>>=
+load('~/upstat-competition/.RData')
+require(psych)
+require(semTools)
+require(mvpart)
+Sys.setenv(TEXINPUTS=getwd(), 
+           BIBINPUTS=getwd(), 
+           BSTINPUTS=getwd())
+@
 
@@
 \begin{frame}{Introduction}
-
+  Is the test psychometrically valid? \pause
+  \textbf{No! It doesn't measure the same thing across groups.}
 \end{frame}
@@
     \begin{itemize}
-
+      \item 10,515 10th grade student test results \pause
+      \item English, Math, and Science components \pause
+      \item Answers from every question for each section \pause
+      \item Basic demographic information from questionnaire \pause
     \end{itemize}
@@
     \begin{frame}{Results}
-
+      <<echo=F,fig=T>>=
+      mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4)
+      @
     \end{frame}

### 2014-04-12 presentation.Rnw
Presentation changes

--- 2014-04-09 presentation.Rnw
@@
 \documentclass{beamer}
+
+\usepackage{apacite}
+\setbeamertemplate{bibliography item}[text]
+
+<<setup, include=F, cache=FALSE, echo=FALSE>>=
+load('~/upstat-competition/.RData')
+require(psych)
+require(semTools)
+require(mvpart)
+setwd("~/upstat-competition/")
+Sys.setenv(TEXINPUTS=getwd(), 
+           BIBINPUTS=getwd(), 
+           BSTINPUTS=getwd())
+@
 
@@
 
-\begin{frame}
+\begin{frame}{Outline}
   \tableofcontents
@@
 \begin{frame}{Introduction}
-
+  \begin{itemize}
+    \item Massachusetts standardized test data \pause
+    \item Is the test psychometrically valid? \pause
+    \item \textbf{No! It doesn't measure the same thing in the same way across groups.}
+  \end{itemize}
 \end{frame}
@@
   \begin{frame}{Measurement Invariance}
-
+     \begin{itemize}
+      \item{Testing:} set of standardized questions designed to elicit responses for use in measuring the traits, capacities, or achievements of an individual.
+       \item{} Tests imply a global solution: the individual should never affect how items on the test perform.
+        \item{} In psychometrics, validity is tested in one of two way: predictive invariance \& measurement (factorial) invariance.
+      \end{itemize}
+  \end{frame}
+  \begin{frame}{Measurement Invariance Process}
+    \begin{enumerate}
+      \item{Configural invariance:} Tests whether the same factor model is found within each subgroup.
+      \item{Weak invariance:} Tests whether in addition to the same factor structure, the same items load equivalently onto the same structure.
+      \item{Strong invariance:} Tests that the intercepts, item loadings, and factor structure are equivalent across groups.
+      \item{Strict invariance:} Tests the assumption that in addition to the preceding steps, the residual variances are equivalent across groups.
+    \end{enumerate}
+  \end{frame}
+  \begin{frame}{Asessing model fit}
+    \begin{itemize}
+      \item{$\chi^2$}\begin{equation} \chi^2_{d.f.}=2(N-1) \cdot F_0 \end{equation}
+      Scales with $N$, making it difficult to fit
+      \item{CFI:} \begin{equation} 1-\frac{\chi^2_{t}-{df_t}}{\chi^2_{n}-{df_n}} \end{equation}
+      \item{NNFI:} \begin{equation} 1-\frac{\frac{\chi^2_{t}}{df_t}}{\frac{\chi^2_{n}}{{df_n}}} \end{equation}
+      \item{RMSEA:} \begin{equation} \sqrt{\frac{\chi^2_{df}-{df_t}}{df_t(N-1)}} \end{equation}
+    \end{itemize}
   \end{frame}
@@
 
-    \begin{frame}{Results}
-
+    \begin{frame}{results}
+      <<fig=T,echo=F>>=
+      fa.diagram(feng,main='English Oblimin')
+      #fa.diagram(fmat,main='Mathematics Oblimin')
+      #fa.diagram(fsci1,main='Biology Oblimin')
+      @
     \end{frame}
 
+    \begin{frame}{Results}
+\vspace{0.25cm}
+\begin{tabular}{|c|c|c|c|c|}
+\hline 
+English & $\chi^2$ & CFI & TLI & RMSEA
+\tabularnewline
+\hline 
+\hline 
+Comparison & -1670.297 & 0.030 & 0.034 & -0.006  \tabularnewline
+\hline 
+Strict fit indices & 6007.770 & 0.918 & 0.912 & 0.030 \tabularnewline
+\hline 
+\end{tabular}
+\end{frame}
+\begin{frame}{Results}
+\vspace{0.25cm}
+\begin{tabular}{|c|c|c|c|c|}
+\hline 
+Mathematics & $\chi^2$ & CFI & TLI & RMSEA
+ \tabularnewline
+\hline 
+\hline 
+Comparison & -998.348 & .014 & .017 & .005  \tabularnewline
+\hline 
+Strict fit indices & 4130.561 & .949 & .946 & .033 \tabularnewline
+\hline 
+\end{tabular}
+\end{frame}
+\begin{frame}{Results}
+\vspace{0.25cm}
+\begin{tabular}{|c|c|c|c|c|}
+\hline 
+Biology & $\chi^2$ & CFI & TLI & RMSEA
+ \tabularnewline
+\hline 
+\hline 
+Comparison & -2132.914 & .026 & .03 & .006  \tabularnewline
+\hline 
+Strict fit indices & 7686.782 & .926 & .922 & .031 \tabularnewline
+\hline 
+\end{tabular}
+\end{frame}
   \subsection{Discussion}
-
     \begin{frame}{Discussion}
+  Strict invariance: what does it mean for these tests?
+      \begin{itemize}
+      \item{SES:}
+      Lack of strict invariance typically indicates that an unmeasured latent variable is differentially affecting the test outcome.
+In other words, the existence of unmeasured effects is specified by the existence of error variance heterogeneity.
+This is despite the invariance of both the factor model, the intercepts, and slopes.
+      \item{Meredith (1993):}
+      When using measures to make decisions “strict factorial invariance is essential.”
+      \end{itemize}
 
@@
   \begin{frame}{Regression Tree}
+  \begin{itemize}
+  \item{} CART allows for complex relationships to be modelled within data without parametric assumptions or linearity.
+  \item{} Multivariate regression trees are ideal in this instance because it can model both the complex relationships, but also non-response, in relation to a multiple vectored response surface: English, Mathematics, \& Science.
+  \end{itemize}
+  \end{frame}
+  \begin{frame}{Results}
+      <<echo=F,fig=T,fig.height=5.5,cache=T>>=
+mvpart(cbind(mrawsc,erawsc,srawsc)~.,data=d4)
+@
+  \end{frame}
+  \subsection{Discussion}
+    \begin{frame}{Discussion}
+    \begin{itemize}
+      \item{Missing data:} Question 3 deals with after college goals.
+These results indicate that students who either intend to go to attain a baccalaureate degree OR skip the question perform better on the three sub-tests.
+      \item{Ethncity:} Amongst students who do plan on a baccalaureate degree, ethnicity efficiently segments the science scores.
+      \item{Quest14:} How often do you use scientific instruments?
+      \item{Confirmation:} Each of the responses were regressed onto multiply imputed forms of the questionnaire, which had equivalent $R^2$, indicating effects were linear only.
+      \end{itemize}
+    \end{frame}
+
+\section{Conclusions}
+  \begin{frame}{Conclusions}
 
@@
 
-  \subsection{Results}
-
-    \begin{frame}{Results}
-
-    \end{frame}
-
-  \subsection{Discussion}
-
-    \begin{frame}{Discussion}
-
+  \begin{frame}[t,allowframebreaks]
+    \frametitle{References}
+    \nocite{*}
+    \bibliography{references}
   \end{frame}

### 2014-04-12 paper-knitr.tex
Presentation RNW

@@
 
-\subsection{Data}
-
 The Massachusetts Comprehensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998.
@@
 
-\subsection{Theoretical structure of the test}
 Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.

### 2014-04-12 presentation.Rnw
Presentation RNW

@@
 \documentclass{beamer}
-
-\usepackage{apacite}
-\setbeamertemplate{bibliography item}[text]
-
-<<setup, include=F, cache=FALSE, echo=FALSE>>=
-load('~/upstat-competition/.RData')
-require(psych)
-require(semTools)
-require(mvpart)
-setwd("~/upstat-competition/")
-Sys.setenv(TEXINPUTS=getwd(), 
-           BIBINPUTS=getwd(), 
-           BSTINPUTS=getwd())
-@
+\usepackage{apacite,geometry,fancyhdr}
 
@@
 \begin{document}
+
+<<setup, include=FALSE, cache=FALSE, echo=FALSE, results='hide'>>=
+library(lavaan)
+library(psych)
+library(mvpart)
+library(semTools)
+
+load("comp.RData")
+
+Sys.setenv(TEXINPUTS=getwd(), 
+           BIBINPUTS=getwd(), 
+           BSTINPUTS=getwd())
+@
 
@@
 
-  \begin{frame}[t,allowframebreaks]
+\begin{frame}[allowframebreaks]
 \frametitle{References}
 \nocite{*}
-    \bibliography{references}
+\bibliography{references}{}
+\bibliographystyle{apacite}
 \end{frame}
-
 \end{document}

### 2014-04-12 test.Rnw
Presentation RNW

--- 2014-04-12 presentation.Rnw
@@
 \documentclass{beamer}
-\usepackage{apacite,geometry,fancyhdr}
+\usepackage{apacite,geometry,fancyhdr,graphicx}
 
@@
 
-    \begin{frame}{results}
-      <<fig=T,echo=F>>=
-      fa.diagram(feng,main='English Oblimin')
-      #fa.diagram(fmat,main='Mathematics Oblimin')
-      #fa.diagram(fsci1,main='Biology Oblimin')
-      @
+    \begin{frame}{Results}
+      \begin{center}
+        \includegraphics[keepaspectratio=true,width=2.7in]{english}
+      \end{center}
+    \end{frame}
+
+    \begin{frame}{Results}
+      \includegraphics[keepaspectratio=true,width=4.3in]{mathematics}
+    \end{frame}
+
+    \begin{frame}{Results}
+      \includegraphics[keepaspectratio=true,width=4.3in]{biology}
     \end{frame}
@@
   \begin{frame}{Conclusions}
-
+    \begin{itemize}
+      \item Fails strict invariance \pause
+      \item Regression tree: split on ethnicity
+    \end{itemize}
   \end{frame}
@@
 \end{frame}
+
+\begin{frame}
+  \frametitle{Appendix A}
+  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
+fa.diagram(feng,main='English Oblimin')
+@
+\end{frame}
+
+\begin{frame}
+  \frametitle{Appendix A}
+  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
+fa.diagram(fmat,main='Mathematics Oblimin')
+@
+\end{frame}
+
+\begin{frame}
+  \frametitle{Appendix A}
+  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
+fa.diagram(fsci1,main='Biology Oblimin')
+@
+\end{frame}
+
 \end{document}
